### 2017-01-20 hw01/hw01.rnw
Homework 1

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 1}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{January 18, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 3,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

library(spdep)
set.seed(-8125)
@

\begin{enumerate}

\item%1
{\it Our text implies and others state outright that the \(BB\), \(BW\), and
\(WW\) statistics reveal pretty much the same thing about spatial correlation.
The \texttt{joincount.mc} function will carry out Monte Carlo tests based on
the \(BB\) and \(WW\) statistics. We do not have an R formula for computing the
\(BW\) statistic but it is possible to carry out a \(BW\) joincount test of
spatial autocorrelation (or clustering) using Geary's \(c\).}

\begin{enumerate}

\item%a
{\it Show the relationship between Geary's \(c\) and \(BW\).}

\begin{align*}
BW&=\frac{1}{2}\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}
(Z(\mathbf{s}_{i})-Z(\mathbf{s}_{j}))^{2} \\
&=S^{2}w_{..}\left(\frac{1}{2S^{2}w_{..}}\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}
(Z(\mathbf{s}_{i})-Z(\mathbf{s}_{j}))^{2}\right) \\
&=S^{2}w_{..}c
\end{align*}

\item%b
{\it Carry out a test based on the \(BW\) statistics using \texttt{geary.mc}.
The data file \texttt{atrplx.dat} will be emailed to you at your math
department email addresses. The first 2 columns contain the spatial coordinates
and the fourth column contains the \(Z\) values you need. Use the R handout to
generate the necessary neighbors and list objects.}

<<prob1b, fig.height = 5>>=
# Read data.
atrplx <- read.table('atrplx.dat', col.names = c('X', 'Y', 'ignore', 'Z'))

# Get rook adjacency list.
atrplx.nb <- dnearneigh(as.matrix(atrplx[, c('X', 'Y')]), d1 = 0, d2 = 1)

# Augment with binary-style w[ij].
atrplx.lw <- nb2listw(atrplx.nb, style = 'B')

# Now do the test.
atrplx.c <- geary.mc(atrplx$Z, atrplx.lw, 999)
atrplx.c

# Make a plot.
hist(atrplx.c$res, breaks = 100, freq = FALSE,
     xlab = 'c', main = 'Permutation Distribution for Geary\'s c')
abline(v = atrplx.c$statistic, lwd = 2)
@

A permutation test with 1,000 permutations results in a p-value of
\Sexpr{atrplx.c$p.value}, which is not convincing evidence the the plants are
spatially clustered.

\item%c
{\it Using the output from \texttt{geary.mc} compute \(BW\) and \(E[BW]\). Do
you expect \(BW < E[BW]\) or \(BW > E[BW]\) in the presence of positive spatial
clustering of the plants? Why or why not?}

<<prob1c>>=
# Sample variance.
var(atrplx$Z)

# Neighbor information.
atrplx.lw

# Mean of the permutation distribution.
mean(atrplx.c$res)
@

We have \(S^{2}=\Sexpr{signif(var(atrplx$Z), 5)}\), and the number of nonzero
links tells us that \(w_{..}=\Sexpr{sum(card(atrplx.nb))}\). So
\begin{equation*}
BW=\Sexpr{signif(var(atrplx$Z), 5)}\times\Sexpr{sum(card(atrplx.nb))}
\times\Sexpr{signif(atrplx.c$statistic, 5)}
=\Sexpr{var(atrplx$Z) * sum(card(atrplx.nb)) * atrplx.c$statistic}.
\end{equation*}
Then the expected value is
\begin{equation*}
BW=\Sexpr{signif(var(atrplx$Z), 5)}\times\Sexpr{sum(card(atrplx.nb))}
\times\Sexpr{signif(mean(atrplx.c$res), 5)}
=\Sexpr{signif(var(atrplx$Z) * sum(card(atrplx.nb)) * mean(atrplx.c$res), 5)}.
\end{equation*}

Under positive spatial clustering I would expect \(BW < E[BW]\) because
clustering will place quadrats with plants together, leading to more
black-black and white-white joins, and fewer black-white joins than would
occur under if plants were placed completely randomly.

\item%d
{\it Reproduce the analysis I presented in class using the \emph{Atriplex}
data. Compare the results of the \(BW\) test to those of the \(BB\) and \(WW\)
test that were discussed in class. Do these statistics all seem to indicate the
same thing about spatial clustering of the plants?}

<<prob1d>>=
atrplx.join <- joincount.mc(factor(atrplx$Z), atrplx.lw, 999)
atrplx.join

par(mfrow = c(1, 2))
hist(atrplx.join[[1]]$res, breaks = 50, freq = FALSE,
     xlab = 'WW', main = 'Permutation Dist. for WW')
abline(v = atrplx.join[[1]]$statistic, lwd = 2)
hist(atrplx.join[[2]]$res, breaks = 50, freq = FALSE,
     xlab = 'BB', main = 'Permutation Dist. for BB')
abline(v = atrplx.join[[2]]$statistic, lwd = 2)
@

The \(WW\) test results in a p-value of \Sexpr{atrplx.join[[1]]$p.value},
no evidence of clustering. However, the \(BB\) test gives a p-value of
\Sexpr{atrplx.join[[1]]$p.value}, strong evidence of clustering. The p-value
of the \(BW\) test is in between these values. These tests do not indicate the
same thing about the distribution of the plants because the \(BB\) statistic
suggests they are clustered, by the \(WW\) and \(BW\) statistics do not.

\item%e
{\it For grins compute Moran's \(I\) and compare that result to those above.}

<<prob1e, fig.height = 5>>=
atrplx.i <- moran.mc(atrplx$Z, atrplx.lw, 999)
atrplx.i

hist(atrplx.i$res, breaks = 50, freq = FALSE,
     xlab = 'I', main = 'Permutation Dist. for Moran\'s I')
abline(v = atrplx.i$statistic, lwd = 2)
@

With 1,000 permutations, we get a p-value of \Sexpr{atrplx.i$p.value}, which
gives weak evidence of spatial clustering. This is not as convincing as the
result of the test based on \(BB\), and definitely inconsisitent with the
tests based on \(WW\) and \(BW\).

\end{enumerate}

\pagebreak
\item%2
{\it Categorize the following examples of spatial data as to their data type:}

\begin{enumerate}

\item%a
{\it Elevations in the foothills of the Allegheny mountains.}

\textbf{These are geostatistical data.}

\item%b
{\it Highest elevation within each state in the United States.}

\textbf{These are lattice data.}

\item%c
{\it Concentration of a mineral in soil.}

\textbf{These are geostatistical data.}

\item%d
{\it Plot yields in a uniformity trial.}

\textbf{These are lattice data.}

\item%e
{\it Crime statistics giving names of subdivisions where break-ins occurred in
the previous year and property loss values.}

If the list only includes the subdivisions where crimes occurred then the list
itself is random, so \textbf{these are marked point process data.}

\item%f
{\it Same as previous, but instead of the subdivisions, the individual
dwelling is identified.}

\textbf{These are marked point process data.}

\item%g
{\it Distribution of oaks and pines in a forest stand.}

\textbf{These are marked point process data.}

\end{enumerate}

\item%3
{\it Show that Moran's \(I\) is a scale-free statistic, i.e. \(Z(s)\) and
\(\lambda Z(s)\) yield the same value for any constant \(\lambda \neq 0\).}

First of all, the mean of \(\lambda Z(\mathbf{s})\) is
\begin{equation*}
\frac{\sum_{i=1}^{n}\lambda Z(\mathbf{s}_{i})}{n}=\lambda\overline{Z}
\end{equation*}
and the sample variance of \(\lambda Z(\mathbf{s})\) is
\begin{align*}
\frac{\sum_{i=1}^{n}(\lambda Z(\mathbf{s}_{i})-\lambda\overline{Z})^{2}}{n-1}
=\lambda^{2}\frac{\sum_{i=1}^{n}(Z(\mathbf{s}_{i})-\overline{Z})^{2}}{n-1}
=\lambda^{2}S^{2}.
\end{align*}
Then Moran's \(I\) for \(\lambda Z(\mathbf{s})\) is
\begin{align*}
\frac{n\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}
(\lambda Z(\mathbf{s}_{i})-\lambda\overline{Z})
(\lambda Z(\mathbf{s}_{j})-\lambda\overline{Z})}
{(n-1)\lambda^{2}S^{2}w_{..}}
&=\frac{n\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}\lambda
(Z(\mathbf{s}_{i})-\overline{Z})\lambda
(Z(\mathbf{s}_{j})-\overline{Z})}
{(n-1)\lambda^{2}S^{2}w_{..}} \\
&=\frac{n\lambda^{2}\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}
(Z(\mathbf{s}_{i})-\overline{Z})
(Z(\mathbf{s}_{j})-\overline{Z})}
{(n-1)\lambda^{2}S^{2}w_{..}} \\
&=\frac{n\sum_{i=1}^{n}\sum_{j=1}^{n}w_{ij}
(Z(\mathbf{s}_{i})-\overline{Z})
(Z(\mathbf{s}_{j})-\overline{Z})}
{(n-1)S^{2}w_{..}} \\
&= I
\end{align*}
which is Moran's \(I\) for \(Z(\mathbf{s})\), so Moran's \(I\) is scale-free.

\item%4
{\it Let \(Y_{1},\dots, Y_{n}\) be normally distributed with unknown mean
\(\mu\) and known variance \(\sigma^{2}\). Let
\(Cov(Y_{i}, Y_{j}) = \sigma^{2}\rho\) for \(i \neq j\). We will further assume
that \(ρ > 0\).}

\begin{enumerate}

\item%a
{\it Show that} \(\displaystyle
Var\left(\overline{Y}\right)=\frac{\sigma^{2}}{n}\left[1+(n-1)\rho\right]\).

Note that there are \(n^{2}-n\) pairs \((i,j)\) such that \(i\neq j\). So,
\begin{align*}
Var\left(\overline{Y}\right)&=Var\left(\frac{\sum_{i=1}^{n}}{n}\right) \\
&=\frac{Var\left(\sum_{i=1}^{n}\right)}{n^{2}} \\
&=\frac{\sum_{i=1}^{n}\sum_{j=1}^{n}Cov(Y_{i},Y_{j})}{n^{2}} \\
&=\frac{\sum_{i=1}^{n}Var(Y_{i})+\sum_{i\neq j}Cov(Y_{i},Y_{j})}{n^{2}} \\
&=\frac{n\sigma^{2}+(n^{2}-n)\sigma^{2}\rho}{n^{2}} \\
&=\frac{\sigma^{2}}{n}[1+(n-1)\rho].
\end{align*}

\item%b
{\it Let \(n = 10\) and \(\rho = 0.26\). Compare and contrast a 95\% confidence
interval for \(\mu\) computed using the true standard deviation of
\(\overline{Y}\) and one computed assuming independence.}

\textbf{Using the true standard deviation:}
\begin{equation*}
\overline{Y}\pm 1.96\frac{\sigma^2}{10}\left[1+9\times 0.26\right]
=\overline{Y}\pm 0.65464\sigma^2
\end{equation*}

\textbf{Assuming independence:}
\begin{equation*}
\overline{Y}\pm 1.96\frac{\sigma^2}{10}
=\overline{Y}\pm 0.196\sigma^2
\end{equation*}

The confidence interval computed using the true standard deviation is quite a
bit wider than the confidence interval computed assuming independence. Ignoring
the correlation in this case means we will think our estimates are more
precise than they actually are.

\item%c
{\it Given independence, we know that \(\overline{Y}\) is the ``best''
estimator of \(\mu\). One nice property it has is that it is a consistent
estimator of the mean. Is \(\overline{Y}\) a consistent estimator of the mean
given the correlation structure above? Justify your answer.}

No, \(\overline{Y}\) is not consistent in this situation. Consistency is
convergence in probability to the quantity of interest, which, by definition
5.5.1 in Casella and Berger, is
\begin{equation*}
\lim_{n\to\infty}P\left(|\overline{Y}-\mu|\geq\epsilon\right)=0
\end{equation*}
for all \(\epsilon>0\). This is not possible because
\begin{equation*}
\lim_{n\to\infty}Var(\overline{Y})
=\lim_{n\to\infty}\frac{\sigma^{2}}{n}[1+(n-1)\rho]
=\frac{\sigma^{2}\rho}{n}
\end{equation*}
is greater than zero if \(\rho>0\).

\item%d
{\it Recall that effective sample size is a measure of the effect of
correlation on inference. An equation for the effective sample size under the
equicorrelation model is}
\begin{equation*}
n'=\frac{n}{1+(n-1)\rho}
\end{equation*}
{\it The effective sample size is defined to be the sample size \(n\) of
uncorrelated observations that provide the same information (in a sense) as a
sample of \(n\) correlated observations.}

\begin{enumerate}

\item%i
{\it Compute the effective sample size when \(n\) = 10, 100, and 1000 and
\(\rho\) = 0.05, 0.1, 0.25, and 0.5.}

<<prob4di, eval = FALSE>>=
n <- matrix(rep(c(10, 100, 1000), 4), ncol = 4,
            dimnames = list(paste('\\(n\\) =', c(10, 100, 1000)),
                            paste('\\(\\rho\\) =', c(0.05, 0.1, 0.25, 0.5))))
rho <- matrix(rep(c(0.05, 0.1, 0.25, 0.5), each = 3), ncol = 4,
              dimnames = list(paste('\\(n\\) =', c(10, 100, 1000)),
                              paste('\\(\\rho\\) =', c(0.05, 0.1, 0.25, 0.5))))
nprime <- n / (1 + (n - 1) * rho)
xtable(nprime, digits = 4, align = '|r|rrrr|')
@
{\centering
<<prob4di, echo = FALSE, results = 'asis'>>=
@
}

\item%ii
{\it Find \(\lim n'\) as \(n \to \infty\).}

Solution:
\begin{align*}
\lim_{n\to\infty}n'&=\lim_{n\to\infty}\frac{n}{1+(n-1)\rho} \\
&=\lim_{n\to\infty}\frac{n}{1+n\rho-\rho} \\
&=\frac{1}{\rho}
\end{align*}

\item%iii
{\it The effect is extreme here but we would not expect to see this type of
correlation structure in a spatial setting. Why not?}

Tobler's First Law of Geography: ``Everything is related to everything else,
but near things are more related than distant things.'' The setting in this
problem ignores distance, making distant locations just as correlated as nearby
locations. This is not how things usually work in reality.

\end{enumerate}

\end{enumerate}

\end{enumerate}

\end{document}