### 2017-01-26 hw02/hw02.rnw
hw2

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 2}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{January 26, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 3,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(6815)
library(spatstat)
data(redwood)
@

\begin{enumerate}

\item%1
{\it It was pointed out in class that, conditional on \(n\) events, event
locations are uniformly distributed for a homogeneous Poisson process. We will
consider the simplest example of this. Consider a one-dimensional process on a
transect of length \(L\), \((0, L]\). Given that one event has occurred on the
interval \((0, L]\) what is the probability that it occurred in the subinterval
\((0, s]\) for \(s < L\)?}

Let \(X\) be the location of the event. Then
\(X|N((0, L]) = 1\sim\mathrm{Unif}(0, L)\). So
\begin{equation*}
P(0 < X \leq s|N((0, L]) = 1) = \int_{0}^{s}\frac{1}{L}ds = \frac{s}{L}.
\end{equation*}

\item%2
{\it We looked at an example of \texttt{quadrat.test} on the amacrine data set
in class. We will use it to analyze another data set, called \texttt{redwood}.
You can read about this in the \texttt{spatstat} help material. You will be
using the \texttt{quadrat.test} function. You can also read about this function
in the help material.}

\begin{enumerate}

\item%a
{\it Read the help pages on the \texttt{quadrat.test} function. What null
hypothesis do they claim to be testing?}

<<prob2a, eval = FALSE>>=
require(spatstat)
data(redwood)
help(redwood) # optional information on the data set
help(quadrat.test)
@

The help file says pretty clearly that ``we test the null hypothesis that the
data pattern is a realisation of Complete Spatial Randomness (the uniform
Poisson point process).''

\item%b
{\it Use \texttt{quadrat.test} on the redwood data set.}

<<prob2b1>>=
redwood.fit <- quadrat.test(redwood)
redwood.fit
@

The quadrat test on a \(5 \times 5\) grid yields a test statistic of
\(\chi^{2}_{\Sexpr{redwood.fit$parameter}}
= \Sexpr{sprintf('%.3f', redwood.fit$statistic)}\) with a p-value of
\(\Sexpr{format.pval(redwood.fit$p.value, 4, 0.0001)}\), very strong evidence
that the locations of the redwood trees do not follow complete spatial
randomness.

{\it The default partitioning of the grid is 5 \(\times\) 5. Does that appear
appropriate here? Justify your answer.}

R gave a warning that some of the grid cells had small expected counts, so the
\(5 \times 5\) grid is too fine for these data. Bigger grid cells would be
better.

\item%c
{\it Redo the analysis using a 3 \(\times\) 3 grid.}

<<prob2c>>=
redwood.fit <- quadrat.test(redwood, nx = 3, ny = 3)
redwood.fit
@

This time we don't get a warning message, so the expected cell counts in the
larger grid cells are big enough to use the \(\chi^{2}\) distribution. The
quadrat test on a \(3 \times 3\) grid yields a test statistic of
\(\chi^{2}_{\Sexpr{redwood.fit$parameter}}
= \Sexpr{sprintf('%.3f', redwood.fit$statistic)}\) with a p-value of
\(\Sexpr{sprintf('%.4f', redwood.fit$p.value)}\), still very strong evidence
that the locations of the redwood trees do not follow complete spatial
randomness.

\item%d
{\it Is the value of the test statistic \(X^{2}\) indicative of clustering,
CSR, or a regular pattern? Justify your answer. Note that I am only asking you
to compare the observed value of \(X^{2}\) to what you would expect under each
of these three patterns. You do not need to calculate a P-value just yet.}

Under the null hypothesis, the test statistic follows a
\(\chi^{2}_{\Sexpr{redwood.fit$parameter}}\) with mean
\Sexpr{redwood.fit$parameter}. The observed value of
\Sexpr{sprintf('%.3f', redwood.fit$statistic)} is larger, indicating greater
than excepted variability in quadrat counts which suggests clustering.

\item%e
{\it The investigator suspected a clustered pattern and the plot would seem to
be consistent with this. Rerun the test with \texttt{alternative="c"} in the
argument list. Give the p-value for the test and interpret the results. Does
the test provide evidence against CSR and for clustering? Justify your answer.}

<<prob2d>>=
redwood.fit <- quadrat.test(redwood, nx = 3, ny = 3, alternative = 'c')
redwood.fit
@

The quadrat test on a \(3 \times 3\) grid with a test statistic of
\(\chi^{2}_{\Sexpr{redwood.fit$parameter}}
= \Sexpr{sprintf('%.3f', redwood.fit$statistic)}\) has a one-sided p-value of
\(\Sexpr{sprintf('%.4f', redwood.fit$p.value)}\), very strong evidence that
the locations of the redwood trees are clustered.

\item%f
{\it We can plot the results of the fit.}

<<prob>>=
par(mar = c(0, 0, 0.5, 0))
plot(redwood.fit)
points(redwood, pch = 16, col = '#00000040') # Add semitransparent points for fun!
@

{\it You will see a plot of the 3 \(\times\) 3 grid. There are 3 numbers in
each cell: the observed count (upper left), expected count under CSR (upper
right), and a scaled residual (lower number). The sum of the scaled residuals
is the \(X^{2}\) statistic. Give the results of the test and using the plot
indicate where CSR seems to break down, if it does.}

The one-sided p-value of \(\Sexpr{sprintf('%.4f', redwood.fit$p.value)}\)
provides very strong evidence that trees are clustered. The upper-left and
center-right cells have fewer trees than expected under CSR, with
large-magnitude negative residuals. The center-left and upper-right cells
have large positive residuals, indicating that these cells contain relatively
dense clusters.

\pagebreak
\item%g
{\it Quadrat size can be important. Repeat the analysis using a 2 \(\times\) 2
grid. Give the results and compare to what we saw with the 3 \(\times\) 3
grid.}

<<prob2g>>=
redwood.fit <- quadrat.test(redwood, nx = 2, ny = 2)
redwood.fit
@

The quadrat test on a \(2 \times 2\) grid yields a test statistic of
\(\chi^{2}_{\Sexpr{redwood.fit$parameter}}
= \Sexpr{sprintf('%.3f', redwood.fit$statistic)}\) with a two-sided p-value of
\(\Sexpr{sprintf('%.4f', redwood.fit$p.value)}\), which is no evidence against
complete spatial randomness and contradicts the results of the test on the
\(3 \times 3\) grid.

\end{enumerate}

\item%3
{\it We will compare results from Monte Carlo procedures based on Poisson
sampling and based on conditioning on the number of observed points. We will
use the \texttt{cells} data set. The R code to accomplish that is shown below.
Compare the two procedures. What do they indicate about the spatial pattern and
why? Which procedure do you like best for this data set and why?}

<<prob3, fig.height = 5.5>>=
data(cells)
hbar <- mean(nndist(cells))
hbar
hbar.pois <- rep(0, 1000)
hbar.cond <- rep(0, 1000)
hbar.pois[1] <- hbar
hbar.cond[1] <- hbar
for(i in 2:1000){
  # Poisson Monte Carlo
  dat.pois <- rpoispp(42)
  hbar.pois[i] <- mean(nndist(dat.pois))
  # Conditional Monte Carlo
  dat.cond <- runifpoint(42)
  hbar.cond[i] <- mean(nndist(dat.cond))
}
par(mfrow = c(2, 1))
hist(hbar.pois, prob = TRUE, xlim = c(0.05, 0.13), main = 'Poisson Monte Carlo')
abline(v = hbar)
hist(hbar.cond, prob = TRUE, xlim = c(0.05, 0.13), main = 'Conditional Monte Carlo')
abline(v = hbar)
# Poisson P-value
2 * sum(hbar.pois >= hbar) / 1000
# Conditional P-value
2 * sum(hbar.cond >= hbar) / 1000
@

Both methods give similar p-values of about 0.002, very strong evidence against
complete spatial randomness. (The observed average nearest neighbor distance
is larger than expected under the null, implying regularity.) The null
distribution for the Poison Monte Carlo method is has a larger spread than the
null distribution for the conditional Monte Carlo method, reflecting
variability in the number of events. For this dataset, the conditional method
is more appropriate because the researcher chose the cells to look at and then
rescaled the viewing window to the unit square. Therefore it is appropriate to
consider permutations of these 42 cells rather than simulating new realizations
with random numbers of cells.

\item%4
{\it Below is the frequency distribution of the number of trees per quadrat in
a sample of 100 quadrats each of radius 6 m.}
\begin{center}\begin{tabular}{c|cccccc}
Trees per quadrat & 0 & 1 & 2 & 3 & 4 & \(\geq 5\) \\
\hline
Count & 34 & 33 & 17 & 7 & 3 & 6
\end{tabular}\end{center}
{\it The data were pooled for counts \(\geq 5\) to meet the assumptions of the
method. Carry out a Poisson goodness-of-fit test based on an assumption of CSR.
Discuss the results. The sample mean of the observed counts was 1.43.}

<<prob4>>=
# Observed number of quadrats in each bin.
Observed <- c(`0` = 34, `1` = 33, `2` = 17, `3` = 7, `4` = 3, `>=5` = 6)

# Expected number of quadrats in each bin.
Expected <- 100 * c(dpois(0:4, 1.43), ppois(4, 1.43, lower.tail = FALSE))

rbind(Observed, Expected)

X2 <- sum((Observed - Expected)^2 / Expected)
X2
pchisq(X2, 4, lower.tail = FALSE)
@

With a test statistic of \(\chi^{2}_{4}
= \Sexpr{sprintf('%.3f', X2)}\) and a p-value of
\Sexpr{sprintf('%.4f', pchisq(X2, 4, lower.tail = FALSE))}
there is very strong evidence that the quadrat counts do not come from a
\(\mathrm{Poisson}(1.43)\) distribution. The test result does not indicate
what type of spatial pattern the data follow, but we observed more counts of
zero and more counts \(\geq\) 5 than expected under CSR, which suggests
clustering.

\pagebreak
\item%5
{\it Suppose we have a realization of a spatial point process consisting of
\(N\) event locations \(\{s_{1}, s_{2},\dots, s_{N}\}\). Let \(H_{i}\) denote
the distance between the \(i\)th event and the nearest neighboring event. The
cumulative distribution function of \(H\) (the nearest event-event distance) is
the \(G\) function. (This problem will be continued on the next homework
assignment).}

\begin{enumerate}

\item%a
{\it What is the \(G\) function if the point process is CSR; i.e. what is
\(G(h) = P(H \leq h)\)?}

\begin{align*}
G(h) &= P(H \leq h) \\
&= P(\text{at least 1 event in a circle of radius }h) \\
&= 1 - P(\text{0 events in a circle of radius }h) \\
&= 1 - \frac{e^{-\lambda \pi h^{2}} \lambda^{0}}{0!} \\
&= 1 - e^{-\lambda \pi h^{2}},\quad h > 0
\end{align*}

\item%b
{\it Find the pdf of \(H\).}

\begin{align*}
f_{H}(h) = \frac{dG(h)}{dh} = 2\lambda \pi h e^{-\lambda \pi h^{2}},
\quad h > 0
\end{align*}

\item%c
{\it Find \(E(H)\) and \(Var(H)\). Hint: you found the pdf but before you start
evaluating a gnarly integral take a close look at that pdf and see if you
cannot identify the family of distributions it belongs to. If you can do that
then you can use that knowledge to find the mean and variance.}

\(H\) follows a Weibull distribution with (in the parameterization of Casella
and Berger) \(\gamma = 2\) and \(\displaystyle\beta = \frac{1}{\lambda\pi}\).
This distribution has mean
\begin{equation*}
E(H) = \beta^{\frac{1}{\gamma}}\Gamma\left(1+\frac{1}{\gamma}\right)
= \frac{1}{\sqrt{\lambda\pi}}\Gamma\left(\frac{3}{2}\right)
= \frac{1}{\sqrt{\lambda\pi}}\frac{1}{2}\Gamma\left(\frac{1}{2}\right)
= \frac{1}{\sqrt{\lambda\pi}}\frac{1}{2}\sqrt{\pi}
= \frac{1}{2\sqrt{\lambda}}
\end{equation*}
and variance
\begin{align*}
Var(H) &= \beta^{\frac{2}{\gamma}}
\left[\Gamma\left(1+\frac{2}{\gamma}\right)
-\Gamma^{2}\left(1+\frac{1}{\gamma}\right)\right] \\
&= \frac{1}{\lambda\pi}\left[\Gamma(2)
-\Gamma^{2}\left(\frac{3}{2}\right)\right] \\
&= \frac{1}{\lambda\pi}\left[1-\left(\frac{\sqrt{\pi}}{2}\right)^{2}\right] \\
&= \frac{1}{\lambda\pi}\left[1-\frac{\pi}{4}\right].
\end{align*}

\end{enumerate}

\end{enumerate}

\end{document}