### 2017-02-10 hw04/hw04.rnw
Homework 4

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 4}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{February 10, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 3,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(2753)
@

\begin{enumerate}

\item%1
{\it For \(\lambda = 30\) generate 9 realizations of CSR on the unit square.
For each realization, construct a kernel estimate of \(\lambda(s)\). How do
the estimated intensity functions compare to the constant intensity under CSR?
What precautions does this exercise suggest with regard to interpreting
estimates of intensity from a single realization (or data set)? The following
R code will simultaneously produce the data and plots.}
<<prob1, fig.height = 4>>=
library(spatstat, quiet = TRUE)
par(mfrow = c(3, 3), mar = c(0, 0, 1, 1), cex = 1)
for(i in 1:9) plot(density(rpoispp(30)), zlim = c(0, 140)) # zlim sets the color scale.
@
The estimated intensity functions do not look constant. Most of them have some
hotspots where the estimated intensity is at least 60, and most also have some
cold spots where there are few points so the estimated intensity is close to
zero. This suggests we shouldn't make much out of our estimated intensity
surface unless we have other information suggesting the process really is
heterogeneous, and even then we should be aware that there may be a lot of
uncertainty around the estimate.

\item%2
{\it In class we looked at a heterogeneous Poisson process on the unit square
with with intensity function}
\begin{equation*}
\lambda(x,y) = \exp(5x + 2y).
\end{equation*}

\begin{enumerate}

\item%a
{\it Simulate a realization of the process using the following R code. Plot
the results and comment.}
<<prob2a, fig.height = 2.5>>=
sim.dat <- rpoispp(function(x, y){exp(5 * x + 2 * y)})
par(mar = c(0, 0, 0.5, 0))
plot(sim.dat)
@
The plot looks about how I expected it to look. There are few events at the
bottom left, many events at the top right, a noticeable increase in event
intensity from left to right, and a subtle increase in intensity from bottom to
top (easily seen at the right edge).

\item%b
{\it Plot simulation envelopes for the \(K\) function (or some suitable
modification of it) and comment.}
<<prob2b>>=
par(mar = c(4, 4, 1, 4))
plot(envelope(sim.dat, Kest, verbose = FALSE), legend = FALSE)
@
The empirical \(K\) function stays far outside of the simulation envelope for
all distances except very small ones. There tend to be more events within a
given distance than expected under CSR, but this is not the same pattern we've
seen for clustered processes, where the \(K\) function is large for small
distances but similar to the CSR \(K\) function for larger distances. This
\(K\) function is close to the CSR \(K\) function at small distances, but moves
away from the value expected under CSR for larger distances because, for most
locations, any region other than a small neighborhood will include points with
higher intensity.

\item%c
{\it Fit a trend model to your data using ppm. Provide me with the parameter
estimates and associated standard errors.}
<<prob2c>>=
sim.fit <- ppm(sim.dat, ~ x + y, correction = 'isotropic')
sim.fit
@
The coefficient estimates, \(\widehat{\beta}_{0} =
\Sexpr{signif(coef(sim.fit)['(Intercept)'], 3)}\) (SE =
\Sexpr{signif(sqrt(vcov(sim.fit)['(Intercept)', '(Intercept)']), 3)}),
\(\widehat{\beta}_{x} = \Sexpr{signif(coef(sim.fit)['x'], 3)}\)
(SE = \Sexpr{signif(sqrt(vcov(sim.fit)['x', 'x']), 3)}) and
\(\widehat{\beta}_{y} = \Sexpr{signif(coef(sim.fit)['y'], 3)}\) (SE =
\Sexpr{signif(sqrt(vcov(sim.fit)['y', 'y']), 3)}), are pretty close to the
true values, with the true values being in the 95\% confidence intervals.

\item%d
{\it Check the fit using \texttt{quadrat.test}. Use
\texttt{method="MonteCarlo"} instead of the large sample chi-squared test.
Plot the results. Discuss.}
<<prob2d, fig.width = 5>>=
sim.qtest <- quadrat.test(sim.fit, method = 'MonteCarlo')
sim.qtest
par(mar = c(0, 0, 1, 0))
plot(sim.qtest)
points(sim.dat, pch = 19, col = '#00000040')
@
The p-value of \Sexpr{sim.qtest$p.value} is small enough to make you think, but
provides at best weak evidence that the model is a poor fit. The only standardized residuals with absolute value \(>\) 2 are in the leftmost column of quadrats
and one right in the center. I wouldn't worry about the left edge because the
expected counts are so close to zero that even 1 event would have a big
residual. In the center, the cluster of points with the standardized residual
of 3 isn't all that unusual because any points in a low-intensity region are
going to stand out. If I encountered these data in the field I would see if a
more complicated surface would fit better, but overall the model fits pretty
well. With 25 quadrats, some of them are bound to be in the tails of the
distribution.

\item%e
{\it Compare these results from this model with those to a model fit under an
assumption of CSR. Summarize the results. Provide me the model comparison
results (AIC comparisons are fine).}
<<prob2e>>=
sim.csr <- ppm(sim.dat, ~ 1, correction = 'isotropic')
sim.csr
AIC(sim.csr)
AIC(sim.fit)
@
The CSR model (intercept-only) assumes a constant intensity, estimated to be
\(\widehat{\lambda} = \exp(\Sexpr{signif(coef(sim.csr), 3)}) =
\Sexpr{signif(exp(coef(sim.csr)), 3)}\) with a 95\%
confidence interval of \(\exp(4.28)=\Sexpr{signif(exp(4.280881), 3)}\) to
\(\exp(4.70)=\Sexpr{signif(exp(4.696392), 3)}\). The heterogeneous model
fits much better, with an improvement in AIC of
\(\Sexpr{signif(AIC(sim.csr)-AIC(sim.fit), 5)}\).

\item%f
{\it Plot a nonparametric estimate of the intensity function. Compare the
fitted surface you got using \texttt{ppm} with the nonparametric (kernel
density estimate) surface in some suitable way.}

<<prob2f, fig.height = 2.5, results = 'hide'>>=
par(mfrow = c(1, 2), mar = c(0, 0, 1, 0))
plot(density(sim.dat))
diagnose.ppm(sim.fit, which = 'smooth')
@
The plot on the left shows a kernel density estimate of the intensity surface,
and the plot on the right shows the smoothed residuals, computed as the fitted
intensity surface minus a kernel density estimate. The residuals have bands
alternating positive and negative, indicating that the model missed some
of the patterns in the data.

\end{enumerate}

\item%3
{\it Recall the use of the \texttt{nncorr} statistic in the Finland Pines data
set. The distribution of heights (the marks) was of interest. We saw that the
nearest neighbor correlation between heights was −0.1839798. We questioned
whether or not this was unusual. Carry out a randomization test to assess
this. You can use the \texttt{rlabel} command to scramble the marks if you
want. Provide me with a histogram of the randomization distribution and a
p-value. Discuss \textbf{BRIEFLY} your results. Provide me with your R-code,
also.}
<<prob3, fig.width = 5.75>>=
finpines.ht <- finpines
marks(finpines.ht) <- marks(finpines)$height
obs_corr <- nncorr(finpines.ht)['correlation']
perm_corr <- c(obs_corr, replicate(999, nncorr(rlabel(finpines.ht))['correlation']))
hist(perm_corr, breaks = 100, freq = FALSE, xlab = 'Correlation',
     main = 'Permutation Distribution of Nearest Neighbor Correlation')
abline(v = obs_corr)
pval <- 2 * mean(perm_corr <= obs_corr)
pval
@
999 permutations yield an approximately symmetric permutation distributon, so
doubling the left-tail p-value results in a two-sided p-value of
\Sexpr{pval} giving no evidence that the heights or nearest neighbors are
correlated.

\item%4
{\it Let's derive a \(K\) function for something other than a CSR process. We
will assume a Neyman-Scott process with the following properties.}
\begin{enumerate}[label=\roman*.]
\item {\it The parent process is a homogeneous Poisson process with intensity
\(\lambda\).}
\item {\it The number of offspring produced by each parent (\(N\)) is
homogeneous Poisson with intensity \(\mu\).}
\item {\it The position of each offspring is determined by a bivariate normal
distribution with mean \((0, 0)\) (i.e. it is centered over the parent) and
variance-covariance matrix \(\sigma^{2}\mathbf{I}\). Note that this implies
that the \(x\) and \(y\) coordinates are determined independently of one
another with the same variance.}
\end{enumerate}
{\it Consider 2 offspring from the same parent located at \((X_{1},Y_{1})\)
and \((X_{2},Y_{2})\).}

\begin{enumerate}

\item%a
{\it What is the distribution of}
\begin{equation*}
W=\frac{(X_{1}-X_{2})^{2}}{2\sigma^{2}}+\frac{(Y_{1}-Y_{2})^{2}}{2\sigma^{2}}?
\end{equation*}
\(X_{1}\), \(X_{2}\), \(Y_{1}\), and \(Y_{2}\) are all independent
\(\mathrm{N}(0, \sigma^{2})\), so \(X_{1}-X_{2}\) and \(Y_{1}-Y_{2}\) are
independent \(\mathrm{N}(0, 2\sigma^{2})\). Then
\(\dfrac{(X_{1}-X_{2})^{2}}{2\sigma^{2}}\) and
\(\dfrac{(Y_{1}-Y_{2})^{2}}{2\sigma^{2}}\) are independent \(\chi^{2}_{1}\).
Therefore, \(W \sim \chi^{2}_{2}\).

\item%b
{\it Note that the Euclidean distance between the 2 points is}
\begin{equation*}
H=\left[(X_{1}-X_{2})^{2}+(Y_{1}-Y_{2})^{2})\right]^{1/2}
=\left(2\sigma^{2}W\right)^{1/2}.
\end{equation*}
{\it Derive the cdf of H.}
\begin{align*}
F(h) &= P(H \leq h) \\
&= P\left(\left(2\sigma^{2}W\right)^{1/2} \leq h\right) \\
&= P\left(W \leq \frac{h^{2}}{2\sigma^{2}}\right) \\
&= \int_{0}^{\frac{h^{2}}{2\sigma^{2}}} \frac{t^{0}e^{-\frac{t}{2}}}
{\Gamma\left(1\right)2^{1}} \mathrm{d}t \\
&= \int_{0}^{\frac{h{2}}{2\sigma^{2}}} \frac{1}{2}e^{-\frac{t}{2}}
\mathrm{d}t \\
&= \left[-e^{-\frac{t}{2}}\right]_{t=0}^{\frac{h^{2}}{2\sigma^{2}}} \\
&= 1 - \exp\left(-\frac{h^{2}}{4\sigma^{2}}\right)
\end{align*}

\item%c
{\it Recall that we were told that the \(K\) function for Neyman-Scott
processes with homogeneous Poisson parent processes and radially symmetric
\(f(h)\) is}
\begin{equation*}
K(h)=\pi h^{2} + \frac{E(N(N-1))}{\lambda E(N)^2}F(h).
\end{equation*}
{\it Use the results from above to find the K function for the described
process.}
\begin{align*}
K(h) &= \pi h^{2} + \frac{E(N(N-1))}{\lambda E(N)^{2}}
\left(1-\exp\left(-\frac{h^{2}}{4}\right)\right) \\
&= \pi h^{2} + \frac{E(N^{2})-E(N)}{\lambda E(N)^{2}}
\left(1-\exp\left(-\frac{h^{2}}{4}\right)\right) \\
&= \pi h^{2} + \frac{Var(N)+E(N)^{2}-E(N)}{\lambda E(N)^{2}}
\left(1-\exp\left(-\frac{h^{2}}{4}\right)\right) \\
&= \pi h^{2} + \frac{\mu+\mu^{2}-\mu}{\lambda\mu^{2}}
\left(1-\exp\left(-\frac{h^{2}}{4}\right)\right) \\
&= \pi h^{2} + \frac{1}{\lambda}
\left(1-\exp\left(-\frac{h^{2}}{4}\right)\right) \\
\end{align*}

\end{enumerate}

\end{enumerate}

\end{document}