### 2017-02-17 hw05/hw05.rnw
Hw5

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 5}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{February 17, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 4,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        xtable.include.rownames = FALSE,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(123)#set.seed(25277)
@

\begin{enumerate}

\item%1
{\it Let \(\gamma(\mathbf{s}_{i}, s_{j}) = \gamma(\mathbf{h}_{ij})\) be a
semivariogram for a second-order stationary spatial process.}

\begin{enumerate}

\item%a
{\it Show that \(\gamma(\mathbf{h}_{ij})=C(\mathbf{0})-C(\mathbf{h}_{ij})\).}
\begin{align*}
\gamma(\mathbf{h}_{ij})
&= \frac{1}{2} Var\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right] \\
&= \frac{1}{2} \left[Var(Z(\mathbf{s}_{i})) + Var(Z(\mathbf{s}_{j}))
- 2 Cov(Z(\mathbf{s}_{i}), Z(\mathbf{s}_{j}))\right] \\
&= \frac{1}{2} Var(Z(\mathbf{s}_{i})) + \frac{1}{2} Var(Z(\mathbf{s}_{j}))
- Cov(Z(\mathbf{s}_{i}), Z(\mathbf{s}_{j})) \\
&= \frac{1}{2} C(\mathbf{0}) + \frac{1}{2} C(\mathbf{0})
- C(\mathbf{s}_{i} - \mathbf{s}_{j}) \\
&= C(\mathbf{0}) - C(\mathbf{h}_{ij})
\end{align*}
\item%b
{\it Show that \(\displaystyle
\sum_{i=1}^{n}\sum_{j=1}^{n} a_{i}a_{j}\gamma(\mathbf{h}_{ij}) \leq 0\)
for any sites \(\mathbf{s}_{i}\), \(i = 1,\dots,n\) and for constants
\(a_{i}\), \(i = 1,\dots,n\) with \(\sum_{i=1}^{n} a_{i} = 0\).}
\begin{align*}
\sum_{i=1}^{n}\sum_{j=1}^{n} a_{i}a_{j}\gamma(\mathbf{h}_{ij})
&= \sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j}[C(\mathbf{0})-C(\mathbf{h}_{ij})] \\
&= C(\mathbf{0})\sum_{i=1}^{n} a_{i} \sum_{j=1}^{n} a_{j}
- \sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j}
 Cov(Z(\mathbf{s}_{i}), Z(\mathbf{s}_{j})) \\
&= 0 - \sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j}
\left[E(Z(\mathbf{s}_{i})Z(\mathbf{s}_{j}))
- E(Z(\mathbf{s}_{i}))E(Z(\mathbf{s}_{j}))\right] \\
&= -\sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j}
E(Z(\mathbf{s}_{i})Z(\mathbf{s}_{j}))
+ \sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j}
E(Z(\mathbf{s}_{i}))E(Z(\mathbf{s}_{j})) \\
&= -E\left(\sum_{i=1}^{n}a_{i}Z(\mathbf{s}_{i})
\sum_{j=1}^{n}a_{j}Z(\mathbf{s}_{j})\right)
+ \sum_{i=1}^{n}\sum_{j=1}^{n}a_{i}a_{j} \mu^{2} \\
&= -E\left(\sum_{i=1}^{n}a_{i}Z(\mathbf{s}_{i})\right)^{2} + 0 \\
&\leq 0
\end{align*}

\end{enumerate}

\item%2
{\it Matheron's semivariogram estimator is}
\begin{equation*}
\widehat{\gamma}(\mathbf{h}) = \frac{1}{2|N(\mathbf{h})|}
\sum_{N(\mathbf{h})} \left\{Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right\}^{2}
\end{equation*}
{\it where \(N(\mathbf{h}) = \{(\mathbf{s}_{i}, \mathbf{s}_{j}):
\mathbf{h} = \mathbf{s}_{i} - \mathbf{s}_{j}\}\) and \(|N(\mathbf{h})|\) is
the number of pairs in the set \(N(\mathbf{h})\).}

\begin{enumerate}

\item%a
{\it Let \(Z(\mathbf{s}) = \mu + e(\mathbf{s})\) where \(E[e(\mathbf{s})]=0\)
with \(\gamma_{Z}(\mathbf{h}) = \gamma_{e}(\mathbf{h})\) (adding a constant to
the random error terms does not change the variance/covariance properties of
the process). Show that \(\widehat{\gamma}(\mathbf{h})\) is unbiased for
\(\gamma_{Z}(\mathbf{h})\). That is, show \(E[\widehat{\gamma}(\mathbf{h})]
= \gamma_{Z}(\mathbf{h})\). Hint: Show that under an assumption of a constant
mean}
\begin{equation*}
\gamma_{Z}(\mathbf{h})
= \frac{1}{2}E\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2}
\end{equation*}

First,
\begin{align*}
\gamma_{Z}(\mathbf{h})
&= \frac{1}{2} Var\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right] \\
&= \frac{1}{2} \left(E\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2}
- \left(E[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})]\right)^{2}\right) \\
&= \frac{1}{2} \left(E\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2}
- \left(E[Z(\mathbf{s}_{i})] - E[Z(\mathbf{s}_{j})]\right)^{2}\right) \\
&= \frac{1}{2} \left(E\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2}
- \left(\mu - \mu\right)^{2}\right) \\
&= \frac{1}{2} E\left[Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2}.
\end{align*}
Then,
\begin{align*}
E\left[\widehat{\gamma}(\mathbf{h})\right]
&= E\left[\frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
\left\{Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right\}^{2}\right] \\
&= \frac{1}{|N(\mathbf{h})|} \sum_{N(\mathbf{h})}\frac{1}{2}E\left[
Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right]^{2} \\
&= \frac{1}{|N(\mathbf{h})|} \sum_{N(\mathbf{h})}\gamma_{Z}(\mathbf{h}_{ij}) \\
&= \frac{1}{|N(\mathbf{h})|} |N(\mathbf{h})|\gamma_{Z}(\mathbf{h}) \\
&= \gamma_{Z}(\mathbf{h})
\end{align*}
so \(\widehat{\gamma}(\mathbf{h})\) is unbiased for \(\gamma_{Z}(\mathbf{h})\).

\pagebreak
\item%b
{\it We pointed out in class that Matheron's Estimator is biased in the
presence of trend. Let \(Z(\mathbf{s}) = \mu(\mathbf{s}) + e(\mathbf{s})\)
with \(E[e(\mathbf{s})] = 0\) and \(\gamma_{Z}(\mathbf{h})
= \gamma_{e}(\mathbf{h})\). Show}
\begin{equation*}
E\left[\widehat{\gamma}(\mathbf{h})\right]
= \gamma_{e}(\mathbf{h}) + \frac{1}{2|N(\mathbf{h})|}
\sum_{N(\mathbf{h})}\left[\mu(\mathbf{s}_{i})-\mu(\mathbf{s}_{j})\right]^{2}.
\end{equation*}

First note that
\begin{equation*}
\gamma_{e}(\mathbf{h})
= \frac{1}{2}E\left[e(\mathbf{s}_{i})-e(\mathbf{s}_{j})\right]^{2}
\end{equation*}
because the mean of \(e(\mathbf{s})\) is constant. Now,
\begin{align*}
E\left[\widehat{\gamma}(\mathbf{h})\right]
&= E\left[\frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
\left\{Z(\mathbf{s}_{i}) - Z(\mathbf{s}_{j})\right\}^{2}\right] \\
&= E\left[\frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
\left\{\mu(\mathbf{s}_{i})+e(\mathbf{s}_{i})
- \mu(\mathbf{s}_{j})-e(\mathbf{s}_{j})\right\}^{2}\right] \\
&= E\left[\frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
\left\{[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]^{2}
+ [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2}
- [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j}))
(e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]\right\}\right] \\
&= \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}\left\{
E[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]^{2}
+ E[\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2}
- E[\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j}))
(e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]\right\} \\
&= \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}\left\{
E[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]^{2}
+ [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2}
- [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]
E[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]\right\} \\
&= \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}\left\{
E[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]^{2}
+ [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2}
- [\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]
(0)\right\} \\
&= \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
E[e(\mathbf{s}_{i}) - e(\mathbf{s}_{j})]^{2}
+ \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
[\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2} \\
&= \gamma_{e}(\mathbf{h})
+ \frac{1}{2|N(\mathbf{h})|} \sum_{N(\mathbf{h})}
[\mu(\mathbf{s}_{i}) - \mu(\mathbf{s}_{j})]^{2}
\end{align*}

\pagebreak
\item%c
{\it Consider the (very) simple model \(Z_{i} = 10 + e_{i}\) where the
\(e_{i}\) are independent normally distributed error terms with variance
\(\sigma^{2} = 81\). We have a pure nugget effect model
\(\gamma_{Z}(\mathbf{h}) = \gamma_{e}(\mathbf{h}) = 81\). Simulate 100
observations of \(Z_{i}\) and calculate the empirical semivariogram assuming
the observations are on a one-dimensional transect.}
<<prob2c>>=
library(geoR)
set.seed(25277)
Zdat <- 10 + rnorm(100, 0, 9)
i <- 1:100
xycoord <- cbind(rep(1, 100), i)
Zvgram <- variog(coords = xycoord, data = Zdat)
par(mar = c(4, 4, 1, 1))
plot(Zvgram)
abline(h = 81)
@
{\it Compare what you see in the plot to the true \(\gamma_{Z}(\mathbf{h})\).
Is the result consistent with part (a) above? Why or why not?}

The empirical semivariogram is nearly horizontal and generally stays near the
true value of \(\gamma(\mathbf{h}) = 81\), which is consistent with the
theoretical unbiasedness shown in part (a). The estimates at different lags
tend to be higher than the true semivariogram, but that is not surprising
because they use the same observed values and are correlated. This is only
one observation of a spatial process; Matheron's estimator is unbiased over
all realizations of the process.

\pagebreak
\item%d
{\it Redo the above calculations based on the (still) simple model
\(Z_{i} = 10 + 10i + e_{i}\), i.e. there is now a linear trend and the process
is no longer stationary.}
<<prob2d>>=
Zdat2 <- 10 + 10 * i + rnorm(100, 0, 9)
Zvgram2 <- variog(coords = xycoord, data = Zdat2)
par(mar = c(4, 4, 1, 1))
plot(Zvgram2)
@
{\it Compare the empirical semivariogram \texttt{Zvgram2} to \text{Zvgram}.
Are the results consistent with part (b) above? Justify your answer.}

The data in (c) came from a stationary process so that semivariogram is roughly
constant. but this empirical semivariogram has a parabolic shape because of
son-stationarity. In this situation, the trend in the mean is linear, with
\(\mu(s_{i})-\mu(s_{j}) = 10(s_{i}-s_{j})\), so the parabolic pattern is
consistent with part (b) which shows that the bias in Matheron's estimator is
quadratic in the difference in means.

\pagebreak
\item%e
{\it Fit a linear model to the data in (d), extract the residuals, and compute
the empirical semivariogram for the residuals. Note that what you are doing is
removing the trend.}
<<prob2e>>=
e.resid <- residuals(lm(Zdat2 ~ i))
evgram <- variog(coords = xycoord, data = e.resid)
par(mar = c(4, 4, 1, 1))
plot(evgram)
abline(h = 81)
@
{\it Compare the 3 empirical semivariograms.}

The residuals have a constant mean of zero, so the empirical semivariogram of
the residuals does not have a pattern of increasing bias like the one seen in
part (d). Like in part (c), the empirical semivariogram is an overestimate but
that is probably because of sampling variability (this is another sample of
size one, after all).

\end{enumerate}

\pagebreak
\item%3
{\it Attached is a data set containing the carbon nitrogen values used in the
carbon/nitrogen data set. The first 2 columns contain the coordinates, total
nitrogen is in the third column, total carbon is in the 4th column and the
ratio is in the last column. We will work with the total carbon data. Use
\texttt{geoR} for the analysis. It will be easiest if you convert the data
into a \texttt{geodata} object as follows.}
<<prob3>>=
CN.dat <- read.table('CN.dat', header = TRUE)
TC.geodata <- as.geodata(CN.dat, coords.col = 1:2, data.col = 4)
@

\begin{enumerate}

\item%a
{\it Calculate the empirical semivariogram. Give initial eyeball estimates of
the nugget effect, sill, and (effective) range.}

<<prob3a>>=
TC.vgram <- variog(TC.geodata)
par(mar = c(4, 4, 1, 1))
plot(TC.vgram)
@

The sill is around 0.015, the effective range looks to be about 150, and the
nugget is probably somewhere near 0.005.

\pagebreak
\item%b
{\it Fit an exponential semivariogram to the carbon data using \(OLS\),
\(WLS\), \(MLE\), and \(REML\) methods. Specify a nugget effect in each case,
i.e. you do not need to consider models without a nugget. Plot the fitted
functions and comment on which one you like best.}

<<prob3b, cache = TRUE, results = 'hide'>>=
tc.e.ols <- variofit(TC.vgram, ini.cov.pars = c(0.015, 150/3), nugget = 0.005,
                     fix.nugget = FALSE, cov.model = 'exponential',
                     weights = 'equal')
tc.e.wls <- variofit(TC.vgram, ini.cov.pars = c(0.015, 150/3), nugget = 0.005,
                     fix.nugget = FALSE, cov.model = 'exponential',
                     weights = 'cressie')
tc.e.ml <- likfit(TC.geodata, ini.cov.pars = c(0.015, 150/3), nugget = 0.005,
                  fix.nugget = FALSE, cov.model = 'exponential',
                  lik.method = 'ML')
tc.e.reml <- likfit(TC.geodata, ini.cov.pars = c(0.015, 150/3), nugget = 0.005,
                    fix.nugget = FALSE, cov.model = 'exponential',
                    lik.method = 'RML')

par(mar = c(4, 4, 1, 1))
plot(TC.vgram)
lines(tc.e.ols, lty = 1)
lines(tc.e.wls, lty = 2)
lines(tc.e.ml, lty = 3)
lines(tc.e.reml, lty = 4)
legend('bottomright', lty = 1:4, legend = c('OLS', 'WLS', 'ML', 'REML'))
@

All of the estimates are similar, but the OLS and WLS estimates look best
because they come closest to \(\widehat{\gamma}\) at the smallest lags. I
would use WLS just because I like the theoretical idea of downweighting the
large lags even though the result is barely distinguishable from the OLS
result here.

\item%c
{\it Redo part (b) by fitting a spherical model.}

<<prob3c, cache = TRUE, results = 'hide'>>=
tc.s.ols <- variofit(TC.vgram, ini.cov.pars = c(0.015, 150), nugget = 0.005,
                     fix.nugget = FALSE, cov.model = 'spherical',
                     weights = 'equal')
tc.s.wls <- variofit(TC.vgram, ini.cov.pars = c(0.015, 150), nugget = 0.005,
                     fix.nugget = FALSE, cov.model = 'spherical',
                     weights = 'cressie')
tc.s.ml <- likfit(TC.geodata, ini.cov.pars = c(0.015, 150), nugget = 0.005,
                  fix.nugget = FALSE, cov.model = 'spherical',
                  lik.method = 'ML')
tc.s.reml <- likfit(TC.geodata, ini.cov.pars = c(0.015, 150), nugget = 0.005,
                    fix.nugget = FALSE, cov.model = 'spherical',
                    lik.method = 'RML')

par(mar = c(4, 4, 1, 1))
plot(TC.vgram, ylim = c(0, 0.022))
lines(tc.s.ols, lty = 1)
lines(tc.s.wls, lty = 2)
lines(tc.s.ml, lty = 3)
lines(tc.s.reml, lty = 4)
legend('bottomright', lty = 1:4, legend = c('OLS', 'WLS', 'ML', 'REML'))
@

Again, I would choose WLS. This time, ML and REML both overestimate the sill,
but WLS and OLS match \(\widehat{\gamma}\) pretty well.

\pagebreak
\item%d
{\it Summarize your results in table format. Compare the results and discuss.}

<<prob3d, eval = FALSE>>=
compare <- do.call(rbind,
                   lapply(list(tc.e.ols, tc.e.wls, tc.e.ml, tc.e.reml,
                               tc.s.ols, tc.s.wls, tc.s.ml, tc.s.reml),
                          function(x){
                            return(data.frame(
                                     Model = x$cov.model,
                                     Method = x$method,
                                     Nugget = x$nugget,
                                     Sill = x$cov.pars[1],
                                     Range = x$cov.pars[2] *
                                       ifelse(x$cov.model == 'exponential', 3, 1)))
                          }))
xtable(compare, digits = 5)
@
\begin{center}
<<prob3d, echo = FALSE, results = 'asis'>>=
@
\end{center}
I multiplied the exponential range parameter by three so that the table
actually shows the effective range. I tried a few different initial range
values, and as long as the initial range was somewhere in the flat section of
the semivariogram, the OLS and WLS range estimate was close to the initial
value (for both models). For a given model, all four estimation methods gave
about the same nugget estimate, though the nugget estimates for the spherical
model were almost 2.5 times as big as the nugget estimates for the exponential
model. OLS and WLS gave similar sill values, and ML and REML also found
similar sill estimates; for the exponential model all four methods found
pretty similar sills, but for the spherical model OLS/WLS and ML/REML yielded
noticeably different sill values. The conclusion I can make from all this is
that fitting a semivariogram model involves many choices that affect the end
result, so you need to check the fitted model against the empirical
semivariogram to see if it looks reasonable. For this dataset, I would use the
exponential model fit by WLS.

\end{enumerate}

\end{enumerate}

\end{document}