### 2017-03-08 hw07/hw07.rnw

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 7}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{March 6, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 4,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(geoR)
library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(218923)
@

\begin{enumerate}

\item%1
{\it A data set (\texttt{wheat.txt}) has been sent to you. The data set
contains yields of wheat recorded at spatial coordinates. Note that the header
is x, y, and z with z being the yields. We will need a couple of different data
object types. Pay attention to the R code below. Do not worry about
anisotropy.}

<<prob1_1>>=
wheat.dat <- read.table('wheat.txt', header = TRUE)
wheat.geodat <- as.geodata(wheat.dat, coords.col = 1:2, data.col = 3)
wheat.grid <- expand.grid(seq(0, 50, l = 25), seq(0, 30, l = 25))
@

\begin{enumerate}

\item%a
{\it Plot the data and comment on the results.}

<<prob1a1>>=
plot(wheat.geodat, breaks = 30)
@

There is a trend in wheat yield, with the lowest yields occuring in the bottom
left of the region and the highest yields in the top right.

\item%b
{\it Produce a plot of the empirical semivariogram of the wheat yields. Can
this plot be trusted for estimation of semivariogram parameters to be used in
kriging. Why or why not?}

<<prob1b1, results = 'hide', fig.height = 5>>=
yield.semivar <- variog(wheat.geodat)
plot(yield.semivar, main = 'Empirical Semivariogram of Yield')
@

No, this plot cannot be trusted because the process is non-stationary so the
empirical semivariogram is biased.

\item%c
{\it We will use the \texttt{surf.ls} function in the spatial library to fit a
quadratic trend model to the yields by ordinary least squares and plot the
empirical semivariogram of the residuals.}
<<prob1c1>>=
library(spatial, quietly = TRUE)
wheat.ls <- surf.ls(2, wheat.dat) # fits a second order polynomial trend surface
resid.dat <- cbind(wheat.dat$x, wheat.dat$y, residuals(wheat.ls))
resid.geodat <- as.geodata(resid.dat, coords.col = 1:2, data.col = 3)
@
{\it Fit an appropriate semivariogram model to the semivariogram using your
method of choice. Justify your final selection.}

Based on the empirical semivariogram (plot on next page) I chose initial
values of 17 for the nugget, 10 for the (effective) range, and 25 for the
total sill (implying a partial sill of 8). I fit exponential and spherical
models by Cressie's WLS and by REML.

<<prob1c2, cache = TRUE, results = 'hide', fig.height = 5>>=
resid.semivar <- variog(resid.geodat)
exp.wls <- variofit(resid.semivar, ini.cov.pars = c(8, 10/3), nugget = 17,
                    fix.nugget = FALSE, cov.model = 'exponential', weights = 'cressie')
exp.reml <- likfit(resid.geodat, ini.cov.pars = c(8, 10/3), nugget = 17,
                   fix.nugget = FALSE, cov.model = 'exponential', lik.method = 'rml')
sph.wls <- variofit(resid.semivar, ini.cov.pars = c(8, 10), nugget = 17,
                    fix.nugget = FALSE, cov.model = 'spherical', weights = 'cressie')
sph.reml <- likfit(resid.geodat, ini.cov.pars = c(8, 10), nugget = 17,
                   fix.nugget = FALSE, cov.model = 'spherical', lik.method = 'rml')
plot(resid.semivar, main = 'Semivariogram of Residuals')
lines(exp.wls, lty = 1)
lines(exp.reml, lty = 2)
lines(sph.wls, lty = 3)
lines(sph.reml, lty = 4)
legend('bottomright', lty = 1:4,
       legend = c('Exponential, WLS', 'Exponential, REML', 'Spherical, WLS', 'Spherical, REML'))
@
<<prob1c3, eval = FALSE>>=
estimates <- rbind(
  `Exponential, WLS` = c(exp.wls$nugget, exp.wls$cov.pars * c(1, 3)),
  `Exponential, REML` = c(exp.reml$nugget, exp.reml$cov.pars * c(1, 3)),
  `Spherical, WLS` = c(sph.wls$nugget, sph.wls$cov.pars),
  `Spherical, REML` = c(sph.reml$nugget, sph.reml$cov.pars)
)
colnames(estimates) <- c('Nugget', 'Partial Sill', 'Range')
xtable(estimates)
@
\begin{center}
<<prob1c3, echo = FALSE, results = 'asis'>>=
@
\end{center}

On the plot, all four estimated models look very similar but the table above
shows some disagreement about the parameters. (Note that I multiplied the
exponential range parameter by 3 so the table shows the effective range for
the exponential models and the range for the spherical models.)

I ultimately chose the spherical model estimated by WLS because WLS does not
depend on normality and the spherical models fit the points at the lower lags
a bit better than the exponential models do.

\item%d
{\it Predict yields using universal kriging and ordinary kriging. Use the
parameter estimates from the residual semivariogram when you do ordinary
kriging. Plot the results along with a plot of the kriging standard errors.
Remember to be careful of that range parameter -- what you enter depends on
which semivariogram model you used.}
<<prob1d, cache = TRUE, dependson = 'prob1c2', results = 'hide', fig.height = 5>>=
# I'll just get the parameters straight out of the semivariogram model object.
wheat.uk <- ksline(wheat.geodat, locations = wheat.grid, cov.model = 'spherical',
                   cov.pars = sph.wls$cov.pars, nugget = sph.wls$nugget,
                   trend = 2, m0 = 'kt')
wheat.ok <- ksline(wheat.geodat, locations = wheat.grid, cov.model = 'spherical',
                   cov.pars = sph.wls$cov.pars, nugget = sph.wls$nugget)

par(mfrow = c(2, 2), mar = c(4, 3, 2, 1))
contour(wheat.uk, main = 'Universal Kriging Predicitons')
contour(wheat.uk, val = sqrt(wheat.uk$krige.var), main = 'Universal Kriging SEs')
contour(wheat.ok, main = 'Ordinary Kriging Predicitons')
contour(wheat.ok, val = sqrt(wheat.ok$krige.var), main = 'Ordinary Kriging SEs')
@
{\it Compare the results and comment.}

The predicted values from both methods are pretty similar, but the prediction
surface from universal kriging smoother than the prediction surface from the
ordinary kriging. The standard errors from universal kriging are larger
because they incorporate uncertainty from estimating the mean surface.

\end{enumerate}

\pagebreak
\item%2
{\it We looked at this example in class. We have a one-dimensional process with
point to point covariance function}
\begin{equation*}
C(s_{i}, s_{j}) = \exp\left(-\frac{3\left|s_{i}-s_{j}\right|}{5}\right).
\end{equation*}
{\it The nugget effect is 0, the sill is 1 and the practical range is 5. The
region \(B\) is defined to be the interval \(B = (2, 4)\) with \(|B| = 2\). The
point to block covariance function is}
\begin{equation*}
\sigma(B, s) = Cov(Z(B), Z(s)) = \frac{1}{2}\int_{2}^{4}
\exp\left(-\frac{3\left|u-s\right|}{5}\right)du.
\end{equation*}

\begin{enumerate}

\item%a
{\it Find the covariance function.}

There are three cases to consider: \(s<2\), \(s>4\), and \(s \in B\).

If \(s<2\) then \(u-s>0\) \(\forall u \in B\), so
\begin{align*}
\frac{1}{2}\int_{2}^{4}\exp\left(-\frac{3|u-s|}{5}\right)du
&= \frac{1}{2}\int_{2}^{4}\exp\left(-\frac{3(u-s)}{5}\right)du \\
&= \frac{1}{2}\left[-\frac{5}{3}\exp\left(-\frac{3(u-s)}
{5}\right)\right]_{u=2}^{4} \\
&= \frac{5}{6} \left[\exp\left(\frac{3}{5}(s-2)\right)
- \exp\left(\frac{3}{5}(s-4)\right)\right].
\end{align*}
If \(s>4\) then \(u-s<0\) \(\forall u \in B\), so
\begin{align*}
\frac{1}{2}\int_{2}^{4}\exp\left(-\frac{3|u-s|}{5}\right)du
&= \frac{1}{2}\int_{2}^{4}\exp\left(\frac{3(u-s)}{5}\right)du \\
&= \frac{1}{2}\left[\frac{5}{3}\exp\left(\frac{3(u-s)}
{5}\right)\right]_{u=2}^{4} \\
&= \frac{5}{6} \left[\exp\left(\frac{3}{5}(4-s)\right)
- \exp\left(\frac{3}{5}(2-s)\right)\right].
\end{align*}
Finally, if \(s \in B\) then
\begin{align*}
\frac{1}{2}\int_{2}^{4}\exp\left(-\frac{3|u-s|}{5}\right)du
&= \frac{1}{2}\int_{2}^{s}\exp\left(-\frac{3|u-s|}{5}\right)du
+ \frac{1}{2}\int_{s}^{4}\exp\left(-\frac{3|u-s|}{5}\right)du \\
&= \frac{1}{2}\int_{2}^{s}\exp\left(\frac{3(u-s)}{5}\right)du
+ \frac{1}{2}\int_{s}^{4}\exp\left(-\frac{3(u-s)}{5}\right)du \\
&= \frac{1}{2}\left[\frac{5}{3}\exp\left(\frac{3(u-s)}
{5}\right)\right]_{u=2}^{s}
+ \frac{1}{2}\left[-\frac{5}{3}\exp\left(-\frac{3(u-s)}
{5}\right)\right]_{u=s}^{4} \\
&= \frac{5}{6} \left[\exp(0) - \exp\left(\frac{3}{5}(2-s)\right)
- \exp\left(\frac{3}{5}(s-4)\right) + \exp(0)\right] \\
&= \frac{5}{6} \left[2 - \exp\left(\frac{3}{5}(2-s)\right)
- \exp\left(\frac{3}{5}(s-4)\right)\right].
\end{align*}
So the point-to-block covariance function is
\begin{equation*}
\sigma(B,s) = \frac{5}{6}\begin{cases}
\exp\left(\frac{3}{5}(s-2)\right) - \exp\left(\frac{3}{5}(s-4)\right) &
s<2 \\
2 - \exp\left(\frac{3}{5}(2-s)\right) - \exp\left(\frac{3}{5}(s-4)\right) &
2 \leq s \leq 4 \\
\exp\left(\frac{3}{5}(4-s)\right) - \exp\left(\frac{3}{5}(2-s)\right) &
s>4
\end{cases}.
\end{equation*}
% \(\sigma(B,3)=0.7519806\) here.
% Numerical integration is the notes got 0.7519603.

\item%b
{\it Find \(\sigma(B, B)\).}

The block-to-block covariance is %variance of the block mean?
\begin{align*}
\sigma(B, B) &= \frac{1}{4}\int_{2}^{4}\int_{2}^{4}\sigma(u,s)duds \\
&= \frac{1}{2}\int_{2}^{4}\sigma(B,s)ds \\
&= \frac{1}{2} \int_{2}^{4} \frac{5}{6}\left(2
- \exp\left(\frac{3}{5}(2-s)\right)
- \exp\left(\frac{3}{5}(s-4)\right)\right)ds \\
&= \frac{5}{12} \left[2s + \frac{5}{3}\exp\left(\frac{3}{5}(2-s)\right)
- \frac{5}{3}\exp\left(\frac{3}{5}(s-4)\right)\right]_{s=2}^{4} \\
&= \frac{5}{12} \left[8 + \frac{5}{3}\exp\left(-\frac{6}{5}\right)
- \frac{5}{3}\exp(0) - 4 - \frac{5}{3}\exp(0)
+ \frac{5}{3}\exp\left(-\frac{6}{5}\right)\right] \\
&= \frac{5}{12} \left[4 - \frac{10}{3}
+ \frac{10}{3}\exp\left(-\frac{6}{5}\right)\right] \\
&= \frac{5}{12} \left[\frac{2}{3}
+ \frac{10}{3}\exp\left(-\frac{6}{5}\right)\right] \\
&= \frac{5}{18} + \frac{25}{18}\exp\left(-\frac{6}{5}\right).
\end{align*}
% \(\sigma(B,B)=0.6961031\) here.
% Numerical integration is the notes got 0.6958757.

\end{enumerate}

\end{enumerate}

\end{document}