### 2017-03-22 project/pres.rnw

\documentclass[unknownkeysallowed]{beamer}
\usepackage{graphicx,epsfig,hyperref,booktabs,lmodern}

\usepackage[backend=bibtex,style=authoryear,citestyle=authoryear-comp]{biblatex}
\addbibresource{references.bib}

\beamertemplatenavigationsymbolsempty

\mode<presentation>
{
  \usetheme{bunsen}
  \setbeamercovered{transparent}
  \setbeamertemplate{items}[circle]
}

% set fonts
%\usepackage{fontspec}
%\setsansfont{Fontin Sans}
%\setbeamerfont{frametitle}{size=\LARGE,series=\bfseries}

% color definitions
\usepackage{color}
\definecolor{uipoppy}{RGB}{225, 64, 5}
\definecolor{uipaleblue}{RGB}{96,123,139}
\definecolor{uiblack}{RGB}{0, 0, 0}

% title slide definition
\title{Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models}
\author{Kenny Flagg}
\institute[MSU]{Department of Mathematical Sciences \\
Montana State University}
\date{March 24, 2017}

\begin{document}

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
opts_chunk$set(echo = FALSE, comment = NA, message = FALSE,
               fig.path = 'pres_figure/', cache.path = 'pres_cache/',
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 2.5,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(pointsize = 10))
knit_theme$set('print')

library(xtable)
options(xtable.table.placement = 'H', width = 80, scipen = 2,
        xtable.sanitize.rownames.function = function(x){return(x)},
        show.signif.stars = FALSE)
@

<<packages, message = FALSE, cache = FALSE>>=
library(spatstat)

set.seed(83534)
@

\setbeamertemplate{background}
{\includegraphics[width=\paperwidth,height=\paperheight]{beamer_title_page1}}
\setbeamertemplate{footline}[default]

\begin{frame}
	\vspace*{.5cm}
	\begin{center}
		\textcolor{white}{\huge Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models \\}
		\vspace{.5cm}
		\textcolor{white}{\Large Kenny Flagg}\\
	\end{center}
\end{frame}

% Set the background for the rest of the slides.
% Insert infoline
\setbeamertemplate{background}
 {\includegraphics[width=\paperwidth,height=\paperheight]{msubg2}}
%\setbeamertemplate{footline}[bunsentheme]

\begin{frame}
\frametitle{Introduction}

\begin{itemize}

\item Goals

\begin{itemize}

\item Estimate inhomogeneous intensity surface from events in a subregion

\item Infer intensity across entire region

\end{itemize}

\item Applications

\begin{itemize}

\item Mapping where endangered species are located

\item Mapping geomegnetic anomalies prior to an unexploded ordnace (UXO)
remediation

\end{itemize}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Maximum Likelihood Intensity Surface Fitting}

\begin{itemize}

\item General point processes

\begin{itemize}

\item The theory is not too complicated but the computation is very difficult

\end{itemize}

\item Poisson processes

\begin{itemize}

\item Doable with numerical methods

\item Log-likelihood of Poisson with intensity \(\lambda(\mathbf{s})\) on
region \(D\) (note typos in \textcite{digglebook})
\begin{align*}
\ell(\lambda) &= \left\{-\mu + n\log(\mu) - \log(n!)\right\}
+ \sum_{i=1}^{n}\left\{\log\left(\lambda(\mathbf{s}_{i})\right)
- \log(\mu)\right\} \\
&= \sum_{i=1}^{n}\log\left(\lambda(\mathbf{s}_{i})\right)
-\int_{D}\lambda(\mathbf{s})d\mathbf{s} - \log(n!).
\end{align*}
where \(\mu=\int_{D}\lambda(\mathbf{s})d\mathbf{s}\)
\end{itemize}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Poisson Process Log-Linear Model}

\begin{itemize}

\item Assuming events are independent (conditional on the intensity function),
\begin{equation*}
\log(\lambda(\mathbf{s})) = \mathbf{x}(\mathbf{s})^{T}\boldsymbol{\beta}
\end{equation*}
where \(\mathbf{x}(\mathbf{s})^{T}\) is a row of predictors at location
\(\mathbf{s}\)

\item Predictors can include covariates, but they must be known across the
whole region

\item \textcite{bermanturner} use dummy points and quadrature to set up an
approximation as a weighted Poisson regression

\item Their method is implemented in \texttt{spatstat}'s \texttt{ppm}, with
\texttt{glm} from base R or \texttt{gam} from \texttt{mgcv} as the back-end

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Simple Example}

<<hw4_sim>>=
simulate_win <- owin(c(0, 1), c(0, 1))
omit_win <- owin(c(0.6, 0.9), c(0.6, 0.9))
observe_win <- setminus.owin(simulate_win, omit_win)
predict_win <- owin(c(-0.5, 1.5), c(-0.5, 1.5))
hw4 <- rpoispp(function(x, y){exp(5 * x + 2 * y)}, win = simulate_win)
hw4_fit_s <- ppm(hw4[observe_win] ~ x + y, correction = 'isotropic')
hw4_pred_s <- predict(hw4_fit_s, window = predict_win, se = TRUE)
@

\begin{itemize}

\item True model

\begin{itemize}

\item Poisson process on the unit square

\item \(\displaystyle \log(\lambda(x,y)) = 5x + 2y\)

\end{itemize}

\item But we don't observe
\(\Sexpr{min(omit_win$xrange)} < x < \Sexpr{max(omit_win$xrange)},
\Sexpr{min(omit_win$yrange)} < y < \Sexpr{max(omit_win$yrange)}\)

\end{itemize}

<<hw4_plot>>=
par(mar = c(0, 0, 1, 0))
plot(hw4, main = 'Event Locations')
plot(omit_win, lty = 2, add = TRUE)
@
\end{frame}

\begin{frame}
\frametitle{Estimated Model}

\begin{itemize}

\item Fit the model
\(\displaystyle \log(\lambda(x,y)) = \beta_{0} + \beta_{1}x + \beta_{2}y\)

\end{itemize}

\begin{minipage}{0.5\textwidth}
<<hw4_fitcoefs_s, results = 'asis'>>=
hw4_coef_s <- summary(hw4_fit_s)$coefs.SE.CI[,c('Estimate', 'S.E.')]
rownames(hw4_coef_s) <- c('\\(\\widehat{\\beta}_{0}\\)',
                          '\\(\\widehat{\\beta}_{1}\\)',
                          '\\(\\widehat{\\beta}_{2}\\)')
xtable(hw4_coef_s)
@
\end{minipage}\begin{minipage}{0.5\textwidth}
<<hw4_dummy_s, fig.width = 3.25>>=
par(mar = c(0, 0, 1, 0))
plot(hw4_fit_s$Q, main = 'Data and Dummy Points')
@
\end{minipage}

\end{frame}

\begin{frame}
\frametitle{Extrapolate the Surface}

\begin{itemize}

\item Use the \texttt{predict} method

\item Specify a new window
\(\Sexpr{min(predict_win$xrange)} < x < \Sexpr{max(predict_win$xrange)},
\Sexpr{min(predict_win$yrange)} < y < \Sexpr{max(predict_win$yrange)}\)

\end{itemize}

<<hw4_resultplot>>=
par(mfrow = c(1, 2), mar = c(0, 0, 3, 0), cex = 1, las = 2)
plot(log(hw4_pred_s$estimate), zlim = c(-4.5, 10.5),
     main = expression(log(hat(lambda)(italic(list(x,y))))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')
plot(log(hw4_pred_s$se), zlim = c(-4.5, 10.5),
     main = expression(log(SE(hat(lambda)(italic(list(x,y)))))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')
@
\end{frame}

\begin{frame}
\frametitle{Where is the Uncertainty?}

\begin{itemize}

\item Relative standard error is lowest where the highest intensity was
observed

\end{itemize}

<<hw4_rse>>=
par(mar = c(0, 0, 1.5, 0), las = 2)
plot(hw4_pred_s$se/hw4_pred_s$estimate, zlim = c(0, 1),
     main = expression(SE(hat(lambda)(italic(list(x,y)))) /
                         hat(lambda)(italic(list(x,y)))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')
@

\end{frame}

\begin{frame}
\frametitle{Simple UXO Site}

<<uxoexampledata>>=
# Define the window for the UXO site.
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
@

\begin{itemize}

\item \Sexpr{sprintf('%.2f', area(site_window) / 43560)} acre region (roughly
\Sexpr{prettyNum(c(-1, 1) %*% site_window$xrange, big.mark = ',')} ft by
\Sexpr{prettyNum(c(-1, 1) %*% site_window$yrange, big.mark = ',')} ft)

\item High density of geomagnetic anomalies around targets

\item Low density of background anomalies

\end{itemize}

<<truth, cache = TRUE>>=
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}
x <- seq(1556880, 1564495, by = 20)
y <- seq(535431, 541130, by = 20)
intense.mat <- matrix(100, nrow = length(x), length(y))
for(i in seq_along(x)){
  for(j in seq_along(y)){
    intense.mat[i, j] <- 100 + gauss.elliptic(x[i], y[j],
        mu.x = 1558400, mu.y = 540000,
        s.a = 800 / (2 * qnorm(0.995)), s.b = 1200 / (2 * qnorm(0.995)),
        r = pi/6, maxrate = 200
      ) + gauss.elliptic(x[i], y[j],
        mu.x = 1562000, mu.y = 537000,
        s.a = 2000 / (2 * qnorm(0.995)), s.b = 900 / (2 * qnorm(0.995)),
        r = 0, maxrate = 200
      )
  }
}
intense.im <- im(t(intense.mat), x, y, unitname = c('foot', 'feet'))

par(mar = c(0, 0, 1, 3))
plot(intense.im, main = 'True Intensity Surface',
     border = NA, zlim = c(100, 300), las = 2)
plot(setminus.owin(owin(c(1556870, 1564495), c(535421, 541130),
                        unitname = c('foot', 'feet')),
     site_window), col = 'white', border = 'white', lwd = 2, add = TRUE)
mtext('Anomalies per Acre', 4, line = -1)
@

\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}

\vspace{-0.5cm}{\tiny
\printbibliography
}
\end{frame}

\end{document}