### 2017-04-01 hw08/hw08.rnw
Homework 8

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 8}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{March 31, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 4,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(gridExtra)
library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(6262)
@

\begin{enumerate}

\item%1
{\it The New York Leukemia data are attached along with homework as a comma
delimited text file. You are to explore how excluding 3 potential outliers
(observations 110, 120, and 121) affects the residual spatial autocorrelation,
regression results, and the conclusions obtained from them. Is weighting still
necessary? Fit \(OLS\), \(WLS\), and appropriate \(GLS\) models following my
steps in the example in class.}

I begin by fitting the OLS model, omitting the outliers, and examining the
semivariogram of the residuals. This semivariogram does not look very
different from the semivariogram of from the model fit the the entire dataset,
so I again use an initial range of 18 and an initial nugget of 0.9. I compare
all six models on the next page.

<<prob1_1>>=
library(nlme)
leukemia <- read.csv('leukemia.csv')

leuk_ols <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121))
plot(Variogram(leuk_ols, maxDist = 120))
@

<<prob1_2>>=
# Fit the weighted and spatial models.
leuk_wls <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121),
                weights = varFixed(~1/pop))
leuk_exp <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121),
                correlation = corExp(c(6, 0.9), form = ~x+y, nugget = TRUE))
leuk_sph <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121),
                correlation = corSpher(c(6, 0.9), form = ~x+y, nugget = TRUE))
leuk_exp_w <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121),
                  correlation = corExp(c(6, 0.9), form = ~x+y, nugget = TRUE),
                  weights = varFixed(~1/pop))
leuk_sph_w <- gls(z ~ pexp + age65 + home, data = leukemia, subset = -c(110, 120, 121),
                  correlation = corSpher(c(6, 0.9), form = ~x+y, nugget = TRUE),
                  weights = varFixed(~1/pop))
@

The table below shows the \texttt{pexp} coefficient estimates and AICs for all
six models. All the slope estimates are similar, but the AICs indicate that
the OLS model has the best fit. With the outliers removed, there is no need
for weighting.

\begin{center}
<<prob1_3, echo = FALSE, results = 'asis'>>=
aic <- AIC(leuk_ols, leuk_wls, leuk_exp, leuk_sph, leuk_exp_w, leuk_sph_w)$AIC

# Recreate the table from the notes.
out_table <- data.frame(rbind(
    summary(leuk_ols)$tTable['pexp', c('Value', 'Std.Error')],
    summary(leuk_wls)$tTable['pexp', c('Value', 'Std.Error')],
    summary(leuk_exp)$tTable['pexp', c('Value', 'Std.Error')],
    summary(leuk_sph)$tTable['pexp', c('Value', 'Std.Error')],
    summary(leuk_exp_w)$tTable['pexp', c('Value', 'Std.Error')],
    summary(leuk_sph_w)$tTable['pexp', c('Value', 'Std.Error')]),
    aic, aic - min(aic),
    row.names = c('OLS', 'WLS', 'Exponential', 'Spherical',
                 'Exponential, Weighted', 'Spherical, Weighted')
  )
colnames(out_table) <- c('\\(\\widehat{\\beta}_{1}\\)',
                         '\\(SE\\left(\\widehat{\\beta}_{1}\\right)\\)',
                         '\\(AIC\\)', '\\(\\Delta AIC\\)')

xtable(out_table, digits = c(0, 4, 4, 3, 3))
@
\end{center}

\item%2
{\it Imagine a lattice process on a \(2 \times 3\) rectangle. The sites
\(\mathbf{s}_{1}\), \(\mathbf{s}_{2}\), and \(\mathbf{s}_{3}\) make up the
first row, the remaining sites make up the second row. Assume that the spatial
connectivity matrix is given by}
\begin{equation*}
\mathbf{W} = \begin{bmatrix}
0 & 1 & 0 & 1 & 0 & 0 \\
1 & 0 & 1 & 0 & 1 & 0 \\
0 & 1 & 0 & 0 & 0 & 1 \\
1 & 0 & 0 & 0 & 1 & 0 \\
0 & 1 & 0 & 1 & 0 & 1 \\
0 & 0 & 1 & 0 & 1 & 0
\end{bmatrix}
\end{equation*}
{\it For a simultaneous and conditional autoregressive scheme with}
\begin{equation*}
\mathrm{Var}\left[\mathbf{Z}(\mathbf{s})\right]
= \sigma^{2} \left(\mathbf{I} - \rho \mathbf{W}\right)^{-1}
\left(\mathbf{I} - \rho \mathbf{W}'\right)^{-1}
\end{equation*}
{\it and}
\begin{equation*}
\mathrm{Var}\left[\mathbf{Z}(\mathbf{s})\right]
= \sigma^{2} \left(\mathbf{I} - \rho \mathbf{W}\right)^{-1}
\end{equation*}
{\it respectively, and with \(\rho = 0.25\) do the following:}

\begin{enumerate}

\item%a
{\it Identify the neighbors of lattice cell \(\mathbf{s}_2\)}.

The neighbors of \(\mathbf{s}_{2}\), corresponding to the columns of
\(\mathbf{W}\) that are nonzero in the second row, are \(\mathbf{s}_{1}\),
\(\mathbf{s}_{3}\), and \(\mathbf{s}_{5}\).

\item%b
{\it Compute the variance-covariance matrices for the SAR and CAR schemes.}

For SAR,

<<prob2b1>>=
W <- matrix(c(0, 1, 0, 1, 0, 0,
              1, 0, 1, 0, 1, 0,
              0, 1, 0, 0, 0, 1,
              1, 0, 0, 0, 1, 0,
              0, 1, 0, 1, 0, 1,
              0, 0, 1, 0, 1, 0),
            nrow = 6)
rho <- 0.25

Sigma_SAR <- solve(diag(6) - rho * W) %*% solve(diag(6) - rho * t(W))
print(Sigma_SAR)
@
so
\begin{equation*}
\mathrm{Var}\left[\mathbf{Z}(\mathbf{s})\right]
= \sigma^{2} \begin{bmatrix}
<<prob2b2, echo = FALSE, results = 'asis'>>=
cat(apply(format(Sigma_SAR, digits = 3), 1, paste, collapse = ' & '),
    sep = ' \\\\\n')
@
\end{bmatrix}.
\end{equation*}

For CAR,

<<prob2b3>>=
Sigma_CAR <- solve(diag(6) - rho * W) 
print(Sigma_CAR)
@
so
\begin{equation*}
\mathrm{Var}\left[\mathbf{Z}(\mathbf{s})\right]
= \sigma^{2} \begin{bmatrix}
<<prob2b4, echo = FALSE, results = 'asis'>>=
cat(apply(format(Sigma_CAR, digits = 3), 1, paste, collapse = ' & '),
    sep = ' \\\\\n')
@
\end{bmatrix}.
\end{equation*}

\pagebreak
\item%c
{\it Determine which of the processes is second-order stationary. Justify
your answer.}

Neither process is second-order stationary because the variance is not
spatially constant. For both processes, the variance of
\(\mathrm{Z}(\mathbf{s}_{2})\) and \(\mathrm{Z}(\mathbf{s}_{5})\) is larger
than the variance in other cells.

\item%d
{\it Describe the correlation patterns that result. Are observations
equicorrelated that are the same distance apart? Do correlations decrease
with increasing lag distance?}

The SAR correlation matrix is
<<prob2d1>>=
D_SAR_invsqrt <- diag(1 / sqrt(diag(Sigma_SAR)))
D_SAR_invsqrt %*% Sigma_SAR %*% D_SAR_invsqrt
@
and the CAR correlation matrix is
<<prob2d2>>=
D_CAR_invsqrt <- diag(1 / sqrt(diag(Sigma_CAR)))
D_CAR_invsqrt %*% Sigma_CAR %*% D_CAR_invsqrt
@
so we see that the correlations generally decrease as the lag increases, but
not all pairs at a given lag are equally correlated. Neighbors within a row
are more highly correlated than neighbors across rows, but pairs of cells in
the same row with lag 2 are less correlated than pairs in different rows with
lag 2.

Assuming square grid cells, the lag distance is equivalent to the Manhattan
distance. The correlation structure appears to reflect the Euclidean distance
because diagonally adjacent cells (lag 2; cells \(\mathbf{s}_{1}\) and
\(\mathbf{s}_{5}\), for example) are more correlated than pairs at lag 2 but
in the same row (e.g., cells \(\mathbf{s}_{1}\) and \(\mathbf{s}_{3}\)), and
the latter have a larger Euclidean distance.

\end{enumerate}

\pagebreak
\item%3
{\it We have 5 binomial count responses at 5 locations, i.e. the number of
successes out of \(n(\mathbf{s}_{i})\) trials. Assume a single covariate
\(X_{1}\) and that there is no overdispersion. Find the diagonal elements of
\(\mathbf{V}_{\mu}^{1/2}\) being sure to express these in terms of
\(\beta_{0}\) and \(\beta_{1}\).}

The model is
\begin{align*}
Z(\mathbf{s}_{i}) &\sim
\mathrm{Binomial}(n(\mathbf{s}_{i}), \mu(\mathbf{s}_{i})); \\
\log\left(\frac{\mu(\mathbf{s}_{i})}{1-\mu(\mathbf{s}_{i})}\right)
&= \beta_{0} + \beta_{1} x_{1}(\mathbf{s}_{i})
\end{align*}
for \(i = 1, \dots, 5\). Then, assuming no overdispersion, the \(i\)th
diagonal element of \(\mathbf{V}_{\mu}^{1/2}\) is
\begin{align*}
\sqrt{v(\mu(\mathbf{s}_{i}))}
&= \sqrt{n(\mathbf{s}_{i})\mu(\mathbf{s}_{i})(1-\mu(\mathbf{s}_{i}))} \\
&= \sqrt{n(\mathbf{s}_{i})
\left(\frac{\exp\left[\beta_{0} + \beta_{1} x_{1}(\mathbf{s}_{i})\right]}
{1 + \exp\left[\beta_{0} + \beta_{1} x_{1}(\mathbf{s}_{i})\right]}\right)
\left(\frac{1}{1 + \exp\left[\beta_{0}
+ \beta_{1} x_{1}(\mathbf{s}_{i})\right]}\right)} \\
&= \frac{\sqrt{n(\mathbf{s}_{i})\exp\left[\beta_{0}
+ \beta_{1} x_{1}(\mathbf{s}_{i})\right]}}
{1 + \exp\left[\beta_{0} + \beta_{1} x_{1}(\mathbf{s}_{i})\right]}.
\end{align*}

\end{enumerate}

\end{document}