### 2017-04-04 project/paper.rnw
About done with the project.

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr,hyperref}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5.5pt}

\usepackage[backend=bibtex,style=authoryear,citestyle=authoryear-comp]{biblatex}
\addbibresource{references.bib}

\newcommand\assignment{Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{April 3, 2017}

\pagestyle{fancy}
\lhead{Stat 534 Project}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = FALSE, comment = NA, message = FALSE,
               fig.path = 'paper_figure/', cache.path = 'paper_cache/',
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 2.5,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.table.placement = 'H', width = 80, scipen = 2,
        xtable.sanitize.rownames.function = function(x){return(x)},
        show.signif.stars = FALSE)
@

<<packages, message = FALSE, cache = FALSE>>=
library(spatstat)

set.seed(83534)
@

\section{Introduction}

One little-studied application of inhomogeneous point process intensity
estimation is the use of an intensity surface estimated from events in a
subregion to predict the intensity over the entire region of interest. This
procedure would be relevant whenver it is known or suspected that some type of
plant, animal, or other item occurs in the region, and the goal of the
analysis is to map the trend in where these tend to be located rather than
estimate parameters of some process at work across hypothetical replicates of
similar regions. It may be prohibitively expensive or difficult to observe all
events over the entire region, so a sample of subregions is taken. For
example, conservationists want to study the spatial distribution of an
endangered plant across a large region of thick jungle so that they can
establish a preserve where the plant is protected. They cannot search the
entire jungle, so they take a simple random sample of quadrats and record the
locations of the plants in those quadrats. They will then want to fit a model
describing the trends in intensity of these plants extrapolated over the
entire jungle. In this setting, the objective is to map the realized spatial
intensity of the plant over this jungle so that these plants can be protected,
not to estimate parameters of the process that arranges plants of this species
at jungles like this one.

Another situation where it is useful to extrapolate point pattern intensity
outside of the observed region is in mapping subsurface geomagnetic anomalies,
such as the munitions debris found at former military test ranges. This is
frequently done in the early stages of an unexploded ordnance (UXO)
remediation, where the intensity surface of inert munitions fragments is
used to identify the locations of targets so that the search for UXO can be
focused on the sections of the site most likely to contain it. The anomalies
are only observed when detection equipment passes directly over them, but the
project leaders are concerened with finding UXO and do not want to waste
resources finding every inert fragment that could be found by metal detectors.
The most common data collection method is to take a systematic sample of
straight-line transects and observe anomalies in rectangular regions centered
along the transects. Frequently, moving averages of the intensity are computed
in circular windows and an intensity map is produced using ordinary kriging to
predict the intensity in a grid of these windows. There are several problems
with this approach: it assumes stationarity when the moving averages are
believed to be non-stationary, it ignores the point process nature of the
data, and it is very sensitive to the window size~\parencite{vspguide,wp}. In
this project, I review the use of polynomial trend surface models for the
log-intensity at a simulated UXO site, ultimately concluding that these
methods are not yet developed well enough to be useful.

\section{Surface Fitting by Maximum Likelihood}

It is theoretically possible to use maximum likelihood or maximum
pseudo-likelihood methods to fit trend surface models or regression models
(with covariates) for the intensity of an inhomogeneous point process, but
these are not widely used because, in the most general cases, the problems
are ``notoriously intractable''~\parencite{digglebook}. However, for spatial
Poisson processes, trend surface models are tenable with the help of some
numerical methods.

The log-likelihood of an unmarked Poisson process on a region \(D\) with
intensity \(\lambda(\mathbf{s})\) is found by conditioning on the number of
events in the following manner. The number of events in \(D\) follows a
Poisson distribution with mean \(\mu=\int_{D}\lambda(\mathbf{s})d\mathbf{s}\).
Given that \(n\) events occured, their locations \(\mathbf{s}_{1},\dots,
\mathbf{s}_{n}\) are independent and identically distributed with density
\(\lambda(\mathbf{s})/\mu\). Then the log-likelihood of the intensity is
\begin{align*}
\ell(\lambda) &= \left\{-\mu + n\log(\mu) - \log(n!)\right\}
+ \sum_{i=1}^{n}\left\{\log\left(\lambda(\mathbf{s}_{i})\right)
- \log(\mu)\right\} \\
&= \sum_{i=1}^{n}\log\left(\lambda(\mathbf{s}_{i})\right)
-\int_{D}\lambda(\mathbf{s})d\mathbf{s} - \log(n!).
\end{align*}
A natural way to model the intensity is to define a log-linear model
\begin{equation*}
\log(\lambda(\mathbf{s})) = \mathbf{x}(\mathbf{s})^{T}\boldsymbol{\beta}
\end{equation*}
where \(\mathbf{x}(\mathbf{s})\) can, in principle, include functions of the
spatial coordinates and also covariates. Unfortunately, the likelihood becomes
\begin{equation*}
\sum_{i=1}^{n}\mathbf{x}(\mathbf{s})^{T}\boldsymbol{\beta}
-\int_{D}\exp\left(\mathbf{x}(\mathbf{s})^{T}
\boldsymbol{\beta}\right)d\mathbf{s} - \log(n!)
\end{equation*}
so \(\mathbf{x}(\mathbf{s})\) must be known (or predicted) either across the
entire region \(D\) or at enough locations to approximate the integral
numerically. Thus, covariates require some extra data collection or modeling
effort to incorporate, but trend surfaces are feasible.

\textcite{bermanturner} proposed a practical method for fitting these models
with generalized linear model (GLM) software using quadratic approximations to
the likelihood, even generalizing to other link functions besides the log link.
Their method is based on partitioning \(D\) such that each subset in the
partition contains at most one event, and then maximizing a weighted
pseudolikelihood based upon a binomial or Poission distribution for the count
of events in each subset. They comment that the covariance matrices produced
by standard GLM software under this method are reasonable approximations to
the correct covariance matrices that would be derived from the true likelihood
as long as the quadrature approximation of the likelihood converges to the
true likelihood as the resolution of the partition is increased.
This was followed up by \textcite{baddeleycoeurjolly14}, who developed a
logistic regression-based approach that avoids the quadratic approximation,
thereby reducing bias in the coefficient estimates. Both methods are available
in the \texttt{spatstat} package~\parencite{spatstat} for R~\parencite{baser}.
The \texttt{spatstat} function \texttt{ppm} implements these methods using the
base R \texttt{glm} function to maximize the likelihood; I will use the Berman
and Turner method because it is the default in \texttt{ppm} and therefore
better reflects the experience of a na\"{i}ve practitioner who is not an
expert on spatial point processes.

\section{True Log-Linear Intensity Surface Example}

<<hw4_sim>>=
simulate_win <- owin(c(0, 1), c(0, 1))
omit_win <- owin(c(0.6, 0.9), c(0.6, 0.9))
observe_win <- setminus.owin(simulate_win, omit_win)
predict_win <- owin(c(-0.5, 1.5), c(-0.5, 1.5))
hw4 <- rpoispp(function(x, y){exp(5 * x + 2 * y)}, win = simulate_win)
@

Before moving on to a realistic example UXO site, I present a simple toy
example using a true log-linear intensity surface to illustrate model fitting
and prediction outside the observed region. Previously this semester, we
worked with an example of a Poisson process on the unit square
\(\{(x,y):\Sexpr{min(simulate_win$xrange)} < x
< \Sexpr{max(simulate_win$xrange)}, \Sexpr{min(simulate_win$yrange)} < y
< \Sexpr{max(simulate_win$yrange)}\}\) with the log-linear intensity surface
\begin{equation*}
\log(\lambda(x,y)) = 5x + 2y.
\end{equation*}
I will continue this example, fitting the model
\begin{equation*}
\log(\lambda(x,y)) = \beta_{0} + \beta_{1}x + \beta_{2}y
\end{equation*}
first using the process observed over the full region (the entire unit
square), and then using the process observed in a subregion, treating
\(\{(x,y): \Sexpr{min(omit_win$xrange)} < x < \Sexpr{max(omit_win$xrange)},
\Sexpr{min(omit_win$yrange)} < y < \Sexpr{max(omit_win$yrange)}\}\) as
unobserved. This unobserved section is a region of relatively high intensity,
so if the estimation or prediction methods are sensitive to the observation
window I expect omitting this section to have a large effect.

I will use both estimated models to predict on
\(\{(x,y): \Sexpr{min(predict_win$xrange)} < x
< \Sexpr{max(predict_win$xrange)}, \Sexpr{min(predict_win$yrange)} < y
< \Sexpr{max(predict_win$yrange)}\}\). Figure~\ref{fig:hw4_plot} shows the
realization of this process that I will use for both model fits.

<<hw4_fit_full>>=
hw4_fit_f <- ppm(hw4 ~ x + y, correction = 'isotropic')
hw4_pred_f <- predict(hw4_fit_f, window = predict_win, se = TRUE)
# By default predictions are on a 128 by 128 grid.
@

<<hw4_fit_s>>=
hw4_fit_s <- ppm(hw4[observe_win] ~ x + y, correction = 'isotropic')
hw4_pred_s <- predict(hw4_fit_s, window = predict_win, se = TRUE)
@

<<hw4_plot, fig.pos = 'b!', fig.cap = 'One realization of a Poisson process with log-linear intensity on the unit square. The left panel shows all of the event locations, with the dashed square outlining of the ``unobserved" region. The right panel shows the events and dummy points used when fitting the model to the subregion; the small square is not considered in constructing the approximation to the likelihood, and the points inside it are discarded.'>>=
par(mfrow = c(1, 2), mar = c(0, 0, 1, 0), cex = 1)
plot(hw4, main = 'Event Locations')
plot(omit_win, lty = 2, add = TRUE)
plot(hw4_fit_s$Q, main = 'Data and Dummy Points')
@

For clarity, I denote the estimated models as
\begin{equation*}
\log\left(\widehat{\lambda}_{f}(x,y)\right) = \widehat{\beta}^{(f)}_{0}
+ \widehat{\beta}^{(f)}_{1}x + \widehat{\beta}^{(f)}_{2}y
\end{equation*}
for the model fit to the full region, and
\begin{equation*}
\log\left(\widehat{\lambda}_{s}(x,y)\right) = \widehat{\beta}^{(s)}_{0}
+ \widehat{\beta}^{(s)}_{1}x + \widehat{\beta}^{(s)}_{2}y
\end{equation*}
for the model fit to the subregion. I use \texttt{ppm} to fit the models
with its default likelihood approximation and isotropic edge correction.
The coefficients and standard errors are similar for both estimated models,
and the true values \(\beta_{0} = 0\), \(\beta_{1} = 5\), and
\(\beta_{2} = 2\) all fall within one standard error of the
estimates~(Tables~\ref{hw4estf} and \ref{hw4ests}).

After estimating the models, I use \texttt{spatstat}'s \texttt{predict.ppm}
method to predict the intensity surface on a \(128 \times 128\) lattice of
locations in the region \(\{(x,y): \Sexpr{min(predict_win$xrange)} < x
< \Sexpr{max(predict_win$xrange)}, \Sexpr{min(predict_win$yrange)} < y
< \Sexpr{max(predict_win$yrange)}\}\)~(Figure~\ref{fig:hw4_fitplot}). As
expected from the similarity of the coefficient estimates, the predicted
log-intensity surfaces for both models show very similar linear trends~(top
row of the figure). At a glance the prediction standard errors for both
models appear similar to the predicted intensities, as seen in the middle
row of the figure~(shown on a logarithmic scale for visibility). This makes
intuitive sense because higher true point process intensity implies greater
variance in the number of points, which should lead to greater uncertainty in
estimation or prediction of the intensity. In a linear model setting we would
expect the prediction standard error to increase with the distance from the
observations, but no such trend is apparent here; even in the bottom left of
the prediction region the standard errors decrease as the predicted intensity
decreases. However, it is illuminating to plot the predictor's relative
standard error (the ratio of the standard error to the predicted intensity,
bottom row of the figure). This ratio is not constant, and in fact it is
lowest where the highest intensity of events was observed.

In summary, the estimation procedure does an adequate job of estimating the
parameters of the true model for these data, and the standard error of the
predicted intensity appears to be more strongly related to the distance from
the observed events than to the distance from the observed region.

<<hw4_fitcoefs_f, results = 'asis'>>=
hw4_coef_f <- summary(hw4_fit_f)$coefs.SE.CI[,c('Estimate', 'S.E.')]
rownames(hw4_coef_f) <- c('\\(\\widehat{\\beta}^{(f)}_{0}\\)',
                          '\\(\\widehat{\\beta}^{(f)}_{1}\\)',
                          '\\(\\widehat{\\beta}^{(f)}_{2}\\)')
print(xtable(hw4_coef_f, label = 'hw4estf',
             caption = 'Estimated coefficients for the log-linear trend model fit using the full region.'),
      table.placement = 't!')
@

<<hw4_fitcoefs_s, results = 'asis'>>=
hw4_coef_s <- summary(hw4_fit_s)$coefs.SE.CI[,c('Estimate', 'S.E.')]
rownames(hw4_coef_s) <- c('\\(\\widehat{\\beta}^{(s)}_{0}\\)',
                          '\\(\\widehat{\\beta}^{(s)}_{1}\\)',
                          '\\(\\widehat{\\beta}^{(s)}_{2}\\)')
print(xtable(hw4_coef_s, label = 'hw4ests',
             caption = 'Estimated coefficients for the log-linear trend model fit using the subregion.'),
      table.placement = 't!')
@

<<hw4_fitplot, fig.pos = 'p', fig.height = 8, fig.cap = 'Estimated log-intensity surface, log-scale prediction standard errors, and relative standard errors from models fit using all events in the full simulation region (left) and events in a subset of the simulation region (right). The regions and event locations are overlaid in white.'>>=
par(mfrow = c(3, 2), mar = c(1, 0, 1.2, 2), las = 2, cex = 1)

plot(log(hw4_pred_f$estimate), zlim = c(-4.5, 10.5),
     main = expression(log(hat(lambda)[f](italic(list(x,y))))))
plot(simulate_win, add = TRUE, border = '#ffffff80')
points(hw4, pch = '.', col = 'white')

plot(log(hw4_pred_s$estimate), zlim = c(-4.5, 10.5),
     main = expression(log(hat(lambda)[s](italic(list(x,y))))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')

plot(log(hw4_pred_f$se), zlim = c(-4.5, 10.5),
     main = expression(log(SE(hat(lambda)[f](italic(list(x,y)))))))
plot(simulate_win, add = TRUE, border = '#ffffff80')
points(hw4, pch = '.', col = 'white')

plot(log(hw4_pred_s$se), zlim = c(-4.5, 10.5),
     main = expression(log(SE(hat(lambda)[s](italic(list(x,y)))))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')

plot(hw4_pred_f$se/hw4_pred_f$estimate, zlim = c(0, 1),
     main = expression(SE(hat(lambda)[f](italic(list(x,y)))) /
                         hat(lambda)[f](italic(list(x,y)))))
plot(simulate_win, add = TRUE, border = '#ffffff80')
points(hw4, pch = '.', col = 'white')

plot(hw4_pred_s$se/hw4_pred_s$estimate, zlim = c(0, 1),
     main = expression(SE(hat(lambda)[s](italic(list(x,y)))) /
                         hat(lambda)[s](italic(list(x,y)))))
plot(observe_win, add = TRUE, border = '#ffffff80')
points(hw4[observe_win], pch = '.', col = 'white')
@

\section{Simulated UXO Site}

<<uxoexampledata, cache = TRUE>>=
# Define the window for the UXO site.
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
@

In my writing project, I evaluated ordinary kriging of moving average
intensity values on a simulated tank training range. I will use the same site
for this project. The site is a
\Sexpr{sprintf('%.2f', area(site_window) / 43560)} acre quadrilateral region
with homogeneous Poisson process producing background anomalies with an
intensity of 100 anomalies per acre, and two regions of concentrated munitions
use generated as Poisson processes with bivariate Gaussian intensity. The
total intesnsity at the center of each concentrated munitions use region is
300 anomalies per acre~(Figure~\ref{fig:truth}).

<<truth, cache = TRUE, dependson = 'uxoexampledata', fig.width = 4, fig.pos = 't', fig.cap = 'True intensity at the simulated UXO site.'>>=
# Bivariate normal intensity function that I originalally wrote for my
# writing project, parameterized in terms of the major and minor axes scales
# and the angle of the major axis from horizontal.
#   a is horizontal axis, b is vertical axis, r is rotation angle
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}

# Create an image of the surface in 20 ft by 20 ft pixels.
x <- seq(1556880, 1564495, by = 20)
y <- seq(535431, 541130, by = 20)
intense.mat <- matrix(100, nrow = length(x), length(y))
for(i in seq_along(x)){
  for(j in seq_along(y)){
    intense.mat[i, j] <- 100 + gauss.elliptic(x[i], y[j],
        mu.x = 1558400, mu.y = 540000,
        s.a = 800 / (2 * qnorm(0.995)), s.b = 1200 / (2 * qnorm(0.995)),
        r = pi/6, maxrate = 200
      ) + gauss.elliptic(x[i], y[j],
        mu.x = 1562000, mu.y = 537000,
        s.a = 2000 / (2 * qnorm(0.995)), s.b = 900 / (2 * qnorm(0.995)),
        r = 0, maxrate = 200
      )
  }
}
intense.im <- im(t(intense.mat), x, y, unitname = c('foot', 'feet'))

par(mar = c(0, 0, 1, 3))
plot(intense.im, main = 'True Intensity Surface',
     border = NA, zlim = c(0, 300), las = 2)
plot(setminus.owin(owin(c(1556870, 1564495), c(535421, 541130),
                        unitname = c('foot', 'feet')),
     site_window), col = 'white', border = 'white', lwd = 2, add = TRUE)
mtext('Anomalies per Acre', 4, line = 2)
@

<<sim2000obs, cache = TRUE, dependson = 'uxoexampledata', fig.pos = 't', fig.cap = 'Observed geomagnetic anomalies from one realization of the UXO site. The transects are not shown on this plot because they would obscure the points.'>>=
# Load the sample data used in my writing project.
# The anomaly file contains the locations of the observed anamolies and the
# cog file contains the course-over-ground waypoints on the transects.
ex_anom <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.anomaly')
ex_path <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.cog')

# Create a spatstat window and ppp object.
ex_win <- intersect.owin(site_window,
  do.call(union.owin, lapply(unique(ex_path$x), function(x){
    return(owin(c(x - 3, x + 3), c(535421, 541130),
           unitname = c('foot', 'feet')))
  })))
ex_ppp <- ppp(ex_anom$x, ex_anom$y, window = ex_win)

# Create a ppp object that uses the full region as the window.
ex_ppp_full <- ppp(ex_anom$x, ex_anom$y, window = site_window)

par(mar = c(0, 0, 1, 0))
plot(ex_ppp, main = 'Observed Geomagnetic Anomalies', pch = '.', border = NA)
plot(site_window, add = TRUE)
@

I simulate data collection along parallel straight-line transects spaced 396
feet apart using a detector with a footprint six feet wide (so there is an
unobserved region 390 feet wide between each pair of adjacent transects). The
location of the first transect is selected randomly, and the locations of all
anomalies within three feet of a transect are
recorded~(Figure~\ref{fig:sim2000obs}). In total,
\Sexpr{sprintf('%.1f', area(ex_win) / 43560)} acres are observed, covering
\Sexpr{sprintf('%.1f\\%%', 100 * area(ex_win) / area(site_window))} of the
region of interest.

If the general structure of the site (such as the number and shape of the
high-intensity regions and the type of background process) is known then the
most appropriate model for the anomaly intensity could be a mixture model or
a trend surface with a particular form, but in practice these details are
rarely available. Therefore, practitioners prefer flexible models without
defined structure. I fit \(p\)-degree polynomial models of the form
\begin{equation*}
\log\left(\lambda\left(x, y\right)\right)
= \sum_{i = 0}^{p} \sum_{j=0}^{p-i} \beta_{ij} x^{i} y^{j}
\end{equation*}
for \(p = 2, 3, \dots, 12\) using 500,000 dummy points randomly selected from
a uniform distribution over the observed region. The coordinates \(x\) and
\(y\) are standardized to have mean 0 and variance 1 in order to reduce
numerical instability for the larger exponents.

<<quad1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))
@

<<polyfit2, cache = TRUE, dependson = 'quad1'>>=
ex_fit2 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 2), correction = 'iso')
@
<<polyfit3, cache = TRUE, dependson = 'quad1'>>=
ex_fit3 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 3), correction = 'iso')
@
<<polyfit4, cache = TRUE, dependson = 'quad1'>>=
ex_fit4 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 4), correction = 'iso')
@
<<polyfit5, cache = TRUE, dependson = 'quad1'>>=
ex_fit5 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 5), correction = 'iso')
@
<<polyfit6, cache = TRUE, dependson = 'quad1'>>=
ex_fit6 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 6), correction = 'iso')
@
<<polyfit7, cache = TRUE, dependson = 'quad1'>>=
ex_fit7 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 7), correction = 'iso')
@
<<polyfit8, cache = TRUE, dependson = 'quad1'>>=
ex_fit8 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 8), correction = 'iso')
@
<<polyfit9, cache = TRUE, dependson = 'quad1'>>=
ex_fit9 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 9), correction = 'iso')
@
<<polyfit10, cache = TRUE, dependson = 'quad1'>>=
ex_fit10 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 10), correction = 'iso')
@
<<polyfit11, cache = TRUE, dependson = 'quad1'>>=
ex_fit11 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 11), correction = 'iso')
@
<<polyfit12, cache = TRUE, dependson = 'quad1'>>=
ex_fit12 <- ppm(ex_Q, ~ polynom(scale(x), scale(y), 12), correction = 'iso')
@

<<polypred1, cache = TRUE, dependson = 'quad1'>>=
# Get predictions and SEs.
ex_p2 <- predict(ex_fit2, window = site_window, se = TRUE)
ex_p3 <- predict(ex_fit3, window = site_window, se = TRUE)
ex_p4 <- predict(ex_fit4, window = site_window, se = TRUE)
ex_p5 <- predict(ex_fit5, window = site_window, se = TRUE)
ex_p6 <- predict(ex_fit6, window = site_window, se = TRUE)
ex_p7 <- predict(ex_fit7, window = site_window, se = TRUE)
ex_p8 <- predict(ex_fit8, window = site_window, se = TRUE)
ex_p9 <- predict(ex_fit9, window = site_window, se = TRUE)
ex_p10 <- predict(ex_fit10, window = site_window, se = TRUE)
ex_p11 <- predict(ex_fit11, window = site_window, se = TRUE)
ex_p12 <- predict(ex_fit12, window = site_window, se = TRUE)

ex_kde <- density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp)
@

Figure~\ref{fig:polyplot1} shows the predicted intensity surfaces for the
polynomial models, along with a kernel density estimate for comparison.
500,000 dummy points do not seem to be enough for the likelihood
approximation to converge to the Poisson likelihood, but the predicted
intensity values are reasonable (discussed further in the next section).
Because of the lack of convergence, I do not do any formal model comparisons.

The second- and third-order polynomial models are too simple to capture the
bimodality of the true surface, but the fourth- and higher-order models have
large, high-intensity regions that coincide with both concentrated munitions
use regions of the true surface. Up until degree nine, the polynomial models
are smoother than the kernel intensity estimate. The ninth-, tenth-, and
eleventh-degree models all look similar to the kernel intensity estimate,
while the twelfth-degree model is noticeably noisier. Edge effects are
noticeable for \(p \geq 4\) with some predictions along the boundary abruptly
dropping to zero or increasing to hundreds of anomalies per acre higher than
any of the predictions in the interior of the region.

The kernel density estimate is computed using \texttt{spatstat}'s default
``simple rule of thumb'' to select the smoothing bandwidth. Because the
anomalies located more than three feet from a transect were not observed, I
would expect the kernel density estimate to show ridges of high intensity
along the transects. As it turns out, the default bandwith is large enough to
smooth across multiple transects so the ridges do not appear. However, the
kernel density estimate is biased low because not all anomalies are observed;
I divide the estimate by the proportion of the area observed as an approximate
bias correction and the result is a surprisingly reasonable intensity map.

<<polyplot1, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = 'Predicted intensity surfaces for eleven polynomial models, and a kernel density estimate for comparision. Note that the 4th- through 12th-degree polynomial models have the same color scale, while the 2nd- and 3rd-degree models and the kernel density estimate use different color scales for visibility.'>>=
par(mfrow = c(4, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, main = '2nd Degree')
plot(43560 * ex_p3$estimate, main = '3rd Degree')
plot(43560 * ex_p4$estimate, zlim = c(0, 300), main = '4th Degree')
plot(43560 * ex_p5$estimate, zlim = c(0, 300), main = '5th Degree')
plot(43560 * ex_p6$estimate, zlim = c(0, 300), main = '6th Degree')
plot(43560 * ex_p7$estimate, zlim = c(0, 300), main = '7th Degree')
plot(43560 * ex_p8$estimate, zlim = c(0, 300), main = '8th Degree')
plot(43560 * ex_p9$estimate, zlim = c(0, 300), main = '9th Degree')
plot(43560 * ex_p10$estimate, zlim = c(0, 300), main = '10th Degree')
plot(43560 * ex_p11$estimate, zlim = c(0, 300), main = '11th Degree')
plot(43560 * ex_p12$estimate, zlim = c(0, 300), main = '12th Degree')
plot(43560 * ex_kde, main = 'Kernel Density')
@

Figure~\ref{fig:polyse1} shows the log-scale standard errors for six of the
polynomial models. For all models except the twelfth-degree model, the
standard error is low across most of the site but high along the boundary. The
standard errors increase with the degree of the model. This is expected and
resonable because locations in the interior of the region have the most nearby
anomalies and dummy points, and predictions from more complex models depend on
more estimated parameters.

If the true number of high-intensity regions was known to be 2, any polynomial
model with \(p \geq 4\) (or the kernel density estimate) would provide an
accurate (if imprecise) indication of where to focus the UXO remediation
effort. However, the plots alone provide little indication of which model has
the ideal level of complexity, but there are several practical difficulties
with fitting these models that should be sorted out before I would be
comfortable making recommendations about how to select an optimal model for
prediction outside the observed subregion.

<<polyse1, fig.width = 6, fig.height = 4, fig.pos = 'p', fig.cap = 'Log standard errors for six of the polynomial models.'>>=
par(mfrow = c(2, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(log(43560 * ex_p8$se), main = '8th Degree\nlog(SE)')
plot(log(43560 * ex_p10$se), main = '10th Degree\nlog(SE)')
plot(log(43560 * ex_p12$se), main = '12th Degree\nlog(SE)')
@

<<dummy, fig.pos = 'p', fig.cap = 'The default dummy points selected by \\texttt{spatstat} (left) and the 500,000 random uniform dummy points that I use to fit the models.'>>=
par(mfrow = c(1, 2), mar = c(1, 0, 1, 0), cex = 1)
plot(ex_fit1$Q$dummy, pch = '.', border = NA,
     main = 'Default Dummy Points')
plot(site_window, add = TRUE)
plot(ex_Q$dummy, pch = '.', border = NA,
     main = 'Manual Dummy Points')
plot(site_window, add = TRUE)
@

\section{Difficulties with Implementation}

I encountered many difficulties in fitting these models that suggest the
methods and software are not yet developed well enough to make them useful for
working with such sparsely sampled data. Most of these difficulties resulted
from the implementation with \texttt{spatstat}, but several were computational
issues.

The troubles with \texttt{spatstat} largely come from how the package sets up
grids of locations for dummy points and predictions. It finds a bounding box
for the region and places grid cells on a lattice across the bounding box, but
its grids are so coarse that most of my transects do not have cells centered
on them. It attempts to create an \(80 \times 80\) grid of dummy points, but
the only useable points are 160 points on two transects; all others are
discarded~(Figure~\ref{fig:dummy}, left). Attempting to fit models in
\texttt{ppm} using these dummy points results in an error.

The \texttt{ppm} function fits models successfully when using the 500,000
dummy points that I draw from a uniform distribution~(Figure~\ref{fig:dummy},
right). However, the \texttt{plot} method gives an error about infinite
values, while the \texttt{predict} method by default uses the same
\(80 \times 80\) grid used for the default dummy points and makes predictions
only in the 160 grid cells that happen to lie inside the observed region.
Fortunately, the \texttt{predict} function behaves when the whole site is
specified as the prediction window.

The biggest disappoinment of using \texttt{spatstat} to fit these models is
that models with splines cannot be plotted. The \texttt{ppm} functions
estimates the parameters without any apparent trouble, but the \texttt{plot}
and \texttt{predict} functions give errors because they do not know how to
interpret the smooth terms in the model formula.

All of the above difficulties are software issues that could be fixed. I also
found two computational limitations that will not be remedied as easily.
First, even after rescaling the coordinates, \(\mathbf{X}^{T}\mathbf{X}\)
cannot be inverted because of numerical instabilities when \(p \geq 18\). Such
models are more complicated than needed for my example site, but it is
important to note that a limit on polynomial model complexity exists.

Second, choosing dummy points is hard. I initially used 1,000 points, but that
was not enough and the intensity values predicted by those models were in the
tens of thousands. It is not apparent that the 500,000 randomly-selected dummy
points I ultimately used are optimal, either. \textcite{wartonshepherd}
discuss dummy point selection when the entire region of interest is observed,
suggesting that the number of dummy points be increased until the maximized
log-likelihood reaches approximate convergence. They use 86,227 points for
their example. I used trial-and-error to increase the number of dummy points,
fitting a couple of my models using 600,000 and then 1,000,000 dummy points;
the maximized log-likelihood changed by thousands. I settled on 500,000
because that was the smallest number I tried that yielded predicted
intensities of the correct magnitude. With that many points, each model takes
about a minute to fit; models with enough dummy points to accurately
approximate the likelihood would likely take much longer to estimate.

\pagebreak
\section{Discussion and Conclusion}

Spatial point process intensity surface models are an intriguing option for
modeling partially-observed inhomogeneous point processes, and polynomial
models can approximate complex surfaces. However, these models are not
yet ready for application to UXO data because the software is not quite there
(yet) and they require heavy computing power. Several avenues for future work
are available, such as seeking more efficient ways to approximate the
likelihood and compute parameter estimates, and further exploring how to
choose dummy points. Then, the next step might be to study whether familiar
model selection tools like AIC or tests of predictive ability are suitable
assuming the likelihood approximation is successful. Presently, it seems
necessary either to observe more of the site area than covered by typical UXO
sampling plans, or work with a model fit that is a rough approximation at
best. Thus the biggest open question might be how much of the region do you
need to observe to trust your model?

\printbibliography

\pagebreak
\appendix
\section*{R Code Appendix}

<<packages, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<hw4_sim, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<hw4_fit_full, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<hw4_fit_s, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<hw4_plot, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<hw4_fitplot, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<uxoexampledata, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<truth, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<sim2000obs, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<quad1, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<polyfit2, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit3, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit4, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit5, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit6, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit7, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit8, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit9, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit10, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit11, echo = TRUE, eval = FALSE>>=
@
\vspace{-25pt}
<<polyfit12, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<polypred1, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<polyplot1, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<polyse1, echo = TRUE, eval = FALSE>>=
@
\vspace{-15pt}
<<dummy, echo = TRUE, eval = FALSE>>=
@

\end{document}