### 2017-04-07 hw09/hw09.rnw
Hw9

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 9}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{April 7, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 4,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(nlme)
library(MASS)
library(geoR)

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(92387)
@

\begin{enumerate}

\item%1
{\it I would like you to try and reproduce some of the results in Table 9.10
on page 395 in Waller and Gotway. They fit several models to the Scottish
lipcancer data set. The ones I am interested in are models \(S+OD\) and the
\(MGLM\). I want you to use \texttt{glmmPQL} to see if you can get close to
their fitted results for \(S+OD\). I provide you with some code for fitting
the nonlinear least squares model to see what you can come up with as an
approximation to their \(MGLM\). We did something similar with the Virginia
lead level data earlier (take a look back at page 50 or thereabouts). I have
spent hours every time I teach this course trying to come close to their
results. So I have two possible outcomes in mind here: (1) you (or some of
you) are able to figure out where I am going wrong which I would love to see,
and/or (2) you get a birds-eye view of the issues related to trying to
incorporate spatial correlation structures into generalized linear models.}

{\it The data set they provided on their website is attached. The last 2
columns are the transformed spatial locations they discuss in the text. If you
look at the SAS code they used they divided these by 1000. I encourage you to
work together on this one. It will go much smoother. I would like for you all
to be ready to discuss this on Friday. I have been a bit easy on the exact
time homeworks are due, accepting several late over the course of the
semester. But I want to be able to talk about this in class on Friday so you
need to have them ready by class time.}

<<data, echo = FALSE>>=
lipcancer <- read.csv('lipcancer.WallerGotway.csv')

# I'm mostly sticking to the same variable names as in their SAS code:
#   http://web1.sph.emory.edu/users/lwaller/book/ch9/scotglms.sas
lipcancer$aff1 <- lipcancer$aff/10
lipcancer$logni <- log(lipcancer$exp)
lipcancer$x <- lipcancer$xcoord/1000
lipcancer$y <- lipcancer$ycoord/1000
lipcancer$logsmr <- log(lipcancer$smr + 1)
lipcancer$expct <- lipcancer$exp
@

I first spent some time trying to understand their semivariogram plot. They
say they computed it from the studentized residuals of a WLS regression of
\(\log(\texttt{SMR}+1)\) against \texttt{aff} using the expected counts as
weights. I translated their SAS code to get the plot on the next page. Two
counties are clear outliers with studentized residuals below \(-2\), and
removing them yields a very different semivariogram. Waller and Gotway don't
give any indication that they give these observations special handling, so
for the purpose of replicating their results I do not omit them.

<<data, eval = FALSE>>=
@
<<wls>>=
# Weighted least squares regression.
stat2 <- lm(logsmr ~ aff1, weights = expct, data = lipcancer)
lipcancer$resids <- rstudent(stat2)

plot(rstudent(stat2) ~ fitted(stat2), main = 'WLS Studentized Residuals vs Fitted Values')
@

Note that the smallest distance in the dataset is
\Sexpr{sprintf('%.1f', min(dist(lipcancer[,c('x', 'y')])))} km, but their
semivariogram has three points at distances smaller than that (the caption
claims that the distances are indeed in km). It would seem like either they
plotted the semivariogram of a different dataset or they rescaled the
distances somehow. The largest distance is
\Sexpr{sprintf('%.1f', max(dist(lipcancer[,c('x', 'y')])))} km,
so I took a wild guess that they estimated the semivariogram up to about half
the largest possible distance, and used units of 10 km for some reason. I
eyeballed the distances on their plot and set manual bins, resulting in the
plot on the next page. The vertical scales are slightly different, but the
shape is very similar. At the top of page 395 they say there is negative
dependence at 72 km to 105 km, but I don't know where they see that.

<<plots, message = FALSE>>=
# Trying to recreate their empirical semivariogram.
resid_geodata <- as.geodata(lipcancer, coords.col = c('x', 'y'), data.col = 'resids')

# uvec specifies the bin centers.
resid_vario <- variog(resid_geodata, uvec = seq(0, 275, 25))
plot(resid_vario, main = 'Empirical Semivariogram')
@

Next, I checked that I understood what I was doing by fitting their
non-spatial models. The R Poisson regression results below match their SAS
results.

<<poisson>>=
# Poisson regression without adjustment for overdispersion.
PR <- glm(observed ~ aff1, family = poisson, offset = logni, data = lipcancer)
summary(PR) # Agrees!
@

\pagebreak
Their Poisson regression with overdispersion results are matched by a
quasiPoisson regression in R, as seen below.

<<quasipoisson>>=
# Poisson regression with adjustment for overdispersion.
PR_OD <- glm(observed ~ aff1, family = quasipoisson, offset = logni, data = lipcancer)
summary(PR_OD) # Agrees!
@

The nonspatial model with uncorrelated random effects and ``adjusted for
overdispersion'' is run-of-the-mill GLMM, because \texttt{glmmPQL} estimates
a within-group residual variance parameter (even though that isn't strictly a
``Poisson'' model because of the overdispersion parameter). The coefficients
and standard errors match those in the book, and squaring the random effect
and residual standard deviations gives values that are off from the variances
in the book by only about 0.02.

<<glmmi, message = FALSE>>=
# Poisson GLMM adjusted for overdispersion.
GLMMI <- glmmPQL(observed ~ aff1 + offset(logni), random = ~1|county,
                 family = poisson, data = lipcancer)
summary(GLMMI) # Coefs/SEs match, variances are close.
@

Waller and Gotway appear to construct non-overdispersed spatial models in SAS
by fixing one of the covariance parameters, but I don't understand their
\texttt{parms} statement. The effect of their \texttt{parms} statement might
be to fix the nugget (within-county variance) at 1, but that is a wild guess.
If that guess is correct, and if SAS uses the nugget as a variance multiplier
(analogous to the quasiPoisson dispersion), then this should have the correct
effect: conditional on the random intercept, the observed count has a Poisson
distribution without overdispersion.

I would construct these models by placing the correlation structure on the
random effects, instead of placing it on the ``residuals'', and estimating the
nugget. Unfortunately, I don't know of any R functions that allow that. So
I'm skipping the non-overdispersion spatial mixed models.

The spatial GLMM with overdispersion is confusing. The nugget should not be
necessary because \texttt{glmmPQL} estimates a residual variance to account
for overdispersion. I initially tried to include both parameters anyway; I
played around with the \texttt{lmeControl} options and got a variety of
cryptic error messages, and never got the optimizer to happily converge.

When leaving out the nugget and using \texttt{optim} to do the optimization,
\texttt{glmmPQL} finds estimates in only 7 reweighting iterations. The BFGS
and L-BFGS algorithms both lead to the same estimates, but they are very
different from the SAS results.

With the \texttt{nlminb} optimizer, \texttt{glmmPQL} gets stuck in a feedback
loop where the working response changes a bit every iteration, so the initial
coefficient values change, then the final coefficient estimates for the
iteration are different from the previous iteration, and so the working
responses change again. One time I ran \texttt{glmmPQL} for 10,000 iterations
and it showed no sign of converging. It's strange that it converges so easily
using \texttt{optim}, so I am suspicious about the \texttt{optim} results and
curious if SAS has convergence issues too.

<<sod>>=
# Conditional spatial GLMM adjusted for overdispersion, spherical covariance.
#corr_S_OD <- corSpher(value = c(263, 0.38), form = ~x+y, nugget = TRUE)
corr_S_OD <- corSpher(value = 263, form = ~x+y)
corr_S_OD <- Initialize(corr_S_OD, lipcancer)
S_OD <- glmmPQL(observed ~ aff1 + offset(logni), random = ~1|county,
                correlation = corr_S_OD, family = poisson,
                control = lmeControl(opt = 'optim', 'BFGS'),
                data = lipcancer)
summary(S_OD) # Way off...
@

Finally, I think the marginal model with overdispersion is conceptually the
easiest to understand of their spatial models. It just allows nearby rates to
be similar to each other, without going to the trouble of predicting a latent
mean for each county. It's a relief than \texttt{gnls} gives the same
estimates as SAS.

<<mglm>>=
# R code for MGLM.
log.model <- function(x1, expct, b0, b1){exp(b0 + log(expct) + b1 * x1)}

corr_MGLM <- corSpher(value = 53, form = ~x+y, nugget = FALSE)
corr_MGLM <- Initialize(corr_MGLM, lipcancer)
MGLM <- gnls(observed ~ log.model(aff1, expct, b0, b1),
             data = lipcancer, start = c(b0 = -0.63, b1 = 0.74),
             correlation = corr_MGLM,
             weights = varPower(form = ~fitted(.), fixed = 0.5))
summary(MGLM) # Agrees!
@

\end{enumerate}

\end{document}