### 2017-04-17 hw10/hw10.rnw
Homework 10

\documentclass[11pt]{article}
\usepackage{fullpage,graphicx,float,amsmath,enumitem,fancyhdr}

\setlist{parsep=5.5pt}
\setlength{\parindent}{0pt}

\newcommand\assignment{Stat 534 Homework 10}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{April 17, 2017}

\pagestyle{fancy}
\lhead{\assignment}
\chead{\duedate}
\rhead{\myname}
\setlength{\headheight}{18pt}
\setlength{\headsep}{2pt}

\title{\assignment}
\author{\myname}
\date{\duedate}

\begin{document}
\maketitle

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 4,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.floating = FALSE,
        xtable.sanitize.rownames.function = function(x) x,
        xtable.sanitize.colnames.function = function(x) x,
        width = 80, scipen = 2, show.signif.stars = FALSE)

set.seed(92387)
@

\begin{enumerate}

\item%1
{\it Empirical Bayes -- NC SIDS data: This is a well-known data set on the
incidence of SIDS (Sudden Infant Death Syndrome) deaths in North Carolina. The
data are available in the \texttt{nc.sids} data set in the \texttt{spdep}
package. We will use emprical Bayes' methods to produce smoothed disease maps
of SIDS incidence in the 1974-78 period. I have been letting you do most of
the coding yourself this semester but getting the plots out for a problem like
this is too much so you can use the R code below. Briefly discuss the results.
In addition to the maps you can look at the raw SMR values and the smoothed
estimates and summarize them in some meaningful way. I am particularly
interested in counties in which the data dominate and counties in which the
empirical Bayes' priors dominate. I will send an R script file with the code
included also. I do not know what all those commands mean myself but I know
they work.}

<<prob1_1>>=
library(maptools)
library(spdep)
library(rgdal)
nc.sids <- readOGR(system.file('etc/shapes/sids.shp', package = 'spdep'), verbose = FALSE)
proj4string(nc.sids) <- CRS('+proj=longlat +ellps=clrk66')
# The IDs are just numbers...
#row.names(nc.sids) <- sapply(slot(nc.sids, 'polygons'), function(x) slot(x, 'ID'))
row.names(nc.sids) <- as.character(nc.sids$NAME) # County names.

rn <- nc.sids$FIPSNO
r <- sum(nc.sids$SID74) / sum(nc.sids$BIR74) 
Expected <- nc.sids$BIR74 * r
SMR <- nc.sids$SID74 / Expected

require(DCluster)
EB.pg <- empbaysmooth(nc.sids$SID74, Expected)$smthrr
EB.LogN <- exp(lognormalEB(nc.sids$SID74, Expected)$smthrr)
EB.Global <- EBest(nc.sids$SID74, Expected)$estmm
ncCR85_nb <- read.gal(system.file('etc/weights/ncCR85.gal',
                                  package = 'spdep' )[1], region.id = rn)
EB.Local <- EBlocal(nc.sids$SID74, Expected, ncCR85_nb)$est

nc <- cbind(nc.sids, SMR, EB.pg, EB.LogN, EB.Global, EB.Local)
names(nc) <- c(names(nc.sids), 'SMR', 'PoissonGamma.EB', 'Lognormal.EB', 'Global.EB', 'Local.EB')
row.names(nc) <- as.character(nc$NAME)
spplot(nc, c('Global.EB', 'PoissonGamma.EB', 'Local.EB', 'Lognormal.EB', 'SMR'))
@

<<prob1_2, eval = FALSE>>=
# Select 15 counties to show.
counties <- sort(c('Anson', 'Dare', sample(as.character(nc$NAME), 13)))
allSMRs <- data.frame(SMR = nc$SMR, `Local EB` = nc$Local.EB,
                      `Lognormal EB` = nc$Lognormal.EB,
                      `Poisson-Gamma EB` = nc$PoissonGamma.EB,
                      `Global EB` = nc$Global.EB,
                      row.names = nc$NAME, check.names = FALSE)
xtable(allSMRs[counties,], digits = 4)
@
<<prob1_2, echo = FALSE, results = 'asis'>>=
@
<<prob1_3, eval = FALSE>>=
meansds <- sapply(allSMRs, function(x){return(c(Min = min(x), Mean = mean(x),
                                                Max = max(x), SD = sd(x)))})
xtable(meansds, digits = 4)
@
\begin{center}
<<prob1_3, echo = FALSE, results = 'asis'>>=
@
\end{center}

All of the smoothing methods clearly get the job done, as all four methods
result in adjecent counties generally having smoothed values that are more
similar than their raw values. The Global method does the most smoothing
(having the lowest standard deviation and the smallest range of smoothed
values), but the Poisson-Gamma results are similar. The Local method does the
least smoothing (biggest SD and range). Unsurprisingly, all the smoothing
methods result in smoothed values with similar means as they all pull the
observed SMRs toward the observed mean.

Anson county is an outlier in raw SMR, and it dominates in the posterior for
each method. For counties with low observed SMRs, the priors tend to dominate
and pull the values toward the mean. The Lognormal method allows large SMRs to
remain large, but still pulls low SMRs upward. The Local method behaves
uniquely, actually smoothing some values \emph{away} from the mean when
neighboring counties had more extreme observed SMRs (e.g. Davidson and Surrey
counties). Local smoothing actually assigns one county (Dare) a smoothed value
of 0 because it and its neighbors all have observed SMRs of zero.

If I knew enough about SIDS to expect spatial autocorrelation, I would use
Local smoothing to accomodate this. Otherwise, I would use either Lognormal or
Global smoothing depending on whether or not I have prior information
suggesting that large observed SMRs are real.

\item%2
{\it Continuing with the NC SIDS data. We assume the Poisson-Gamma model.}

\begin{enumerate}

\item%a
{\it Find the posterior distribution assuming a more informative Gamma prior
with \(\alpha = \beta = 4\).}

We assume \(Y_{i}|\theta \sim \mathrm{Poisson}(E_{i}\theta)\),
\(\theta \sim \mathrm{Gamma}(4, 4)\). We observe
\(\sum_{i=1}^{n}y_{i} = \Sexpr{sum(nc$SID74)}\) SIDS cases in 1974 across all
counties. I define the exposure to be the number of births in thousands, so
the response and the exposure are on similar scales. Thus
\(\sum_{i=1}^{n}E_{i} = \Sexpr{sum(nc$BIR74) / 1000}\).

Then the posterior distribution is \(\theta|\mathbf{y} \sim \mathrm{Gamma}(
\Sexpr{sum(nc$SID74) + 4},\, \Sexpr{sum(nc$BIR74) / 1000 + 4})\).

<<prob2a1, eval = FALSE, echo = FALSE>>=
hist(nc$SID74, breaks = 20, freq = FALSE, ylim = c(0, 1),
     xlab = 'SIDS Cases', main = 'Distribution of SIDS Counts')
curve(dgamma(x, 4, 4), n = 501, lty = 2, add = TRUE)
curve(dgamma(x, sum(nc$SID74) + 4, sum(nc$BIR74) / 1000 + 4), n = 501, lty = 1, add = TRUE)
legend('topright', lty = c(2, 1), legend = c('Prior', 'Posterior'))
@

\item%b
{\it Find the posterior distribution assuming a non-informative Gamma prior
with \(\alpha = \beta = 0.1\).}

The posterior distribution is \(\theta|\mathbf{y} \sim \mathrm{Gamma}(
\Sexpr{sum(nc$SID74) + 0.1}, \Sexpr{sum(nc$BIR74) / 1000 + 0.1})\).

<<prob2a2, eval = FALSE, echo = FALSE>>=
hist(nc$SID74, breaks = 20, freq = FALSE, ylim = c(0, 1),
     xlab = 'SIDS Cases', main = 'Distribution of SIDS Counts')
curve(dgamma(x, 0.1, 0.1), n = 501, from = 0.1, lty = 2, add = TRUE)
curve(dgamma(x, sum(nc$SID74) + 0.1, sum(nc$BIR74) / 1000 + 0.1), n = 501, lty = 1, add = TRUE)
legend('topright', lty = c(2, 1), legend = c('Prior', 'Posterior'))
@

\pagebreak
\item%c
{\it Summarize the role of the data and the prior in the posterior
distributions by comparing the observed SMR and the posterior means from the
2 Bayesian approaches. A plot would help and you should be able to modify the
code above to produce those plots. Just add the vectors of the poterior
distribution to \texttt{nc.sids} using \texttt{cbind} and then use
\texttt{spplot} to get plots of SMR, and the pure Bayes' posterior means.}

Both prior distributions have mean 1, and the observed SIDS counts have mean
\Sexpr{mean(nc$SID74)}. The informative prior results in a posterior mean of
\(\Sexpr{sum(nc$SID74) + 4} / \Sexpr{sum(nc$BIR74) / 1000 + 4}
= \Sexpr{sprintf('%.3f', (sum(nc$SID74) + 4) / (sum(nc$BIR74) / 1000 + 4))}\)
and the non-informative prior yields a posterior mean of
\(\Sexpr{sum(nc$SID74) + 0.1} / \Sexpr{sum(nc$BIR74) / 1000 + 0.1} =
\Sexpr{sprintf('%.3f', (sum(nc$SID74) + 0.1) / (sum(nc$BIR74) / 1000 + 0.1))}\).
The informative prior pulls the posterior mean toward 1, but the data dominate
over the non-informative prior resulting in a posterior mean close to the
data mean.

To make a smoothed spatial map, I use the hierarchical model where
\(Y_{i}|\gamma_{i} \sim \mathrm{Poisson}(E_{i}\gamma_{i})\) and
\(\gamma_{i} \sim \mathrm{Gamma}(\alpha, \beta)\). Then the posterior
distribution for each county's mean is
\(\gamma_{i}|y_{i} \sim \mathrm{Gamma}(\alpha + y_{i}, \beta + E_{i})\). The
result is that the informative prior dominates (doing a lot of smoothing)
while the non-informative prior allows the data to dominate (resulting in a
map that is barely distinguishable from map of the data). To create maps on
SMR scale, I multiplied the posterior means by the exposure and divided by the
expected counts.

<<prob2c2>>=
nc.bayes <- cbind(nc,
                  (nc$BIR74 / 1000) * (nc$SID74 + 0.1) / (nc$BIR74 / 1000 + 0.1) / Expected,
                  (nc$BIR74 / 1000) * (nc$SID74 + 4) / (nc$BIR74 / 1000 + 4) / Expected)
names(nc.bayes) <- c(names(nc), 'Non.Informative', 'Informative')
spplot(nc.bayes, c('Non.Informative', 'Informative', 'SMR'))
@

\end{enumerate}

\pagebreak
\item%3
{\it See the notes for the details on the Global Estimator (page 85). Marshall
assumed, with \(E_{i}\) being the expected count that}
\begin{equation*}
Z_{i}|\gamma_{i} \sim \mathrm{Poi}(E_{i}\gamma_{i})
\end{equation*}
{\it but made no assumption about the distribution of the spatially varying
relative risks or SMR values \(\gamma_{i}\). We denote the estimates of SMR by
\(r_{i} = Z_{i} / E_{i}\). We make no distributional assumptions about the
\(\gamma_{i}\) values but do assume that the prior means and variances exist.
Denote them \(m_{\gamma_{i}}\) and \(v_{\gamma_{i}}\), respectively. Marshall
found method-of-moments estimators of the marginal or unconditional mean and
variance of the \(\gamma_{i}\).}

\begin{enumerate}

\item%a
{\it Find the conditional mean and variance of \(r_{i}\). That is find
\(E(r_{i}|\gamma_{i})\) and \(Var(r_{i}|\gamma_{i})\).}

\vspace{-11pt}
\begin{equation*}
E(r_{i}|\gamma_{i}) = E\left(\left.\frac{Z_{i}}{E_{i}}\right|\gamma_{i}\right)
= \frac{E(Z_{i}|\gamma_{i})}{E_{i}} = \frac{E_{i}\gamma_{i}}{E_{i}}
= \gamma_{i}
\end{equation*}
\begin{equation*}
Var(r_{i}|\gamma_{i})
= Var\left(\left.\frac{Z_{i}}{E_{i}}\right|\gamma_{i}\right)
= \frac{Var(Z_{i}|\gamma_{i})}{E_{i}^{2}}
= \frac{E_{i}\gamma_{i}}{E_{i}^{2}}
= \frac{\gamma_{i}}{E_{i}}
\end{equation*}

\item%b
{\it Show that the marginal mean of \(r_{i}\) is equal to the prior mean of
\(\gamma_{i}\).}

\vspace{-11pt}
\begin{align*}
E(r_{i}) &= E(E(r_{i}|\gamma_{i})) \\
&= \int_{0}^{\infty} E(r_{i}|\gamma_{i}) f(\gamma_{i}) d\gamma_{i} \\
&= \int_{0}^{\infty} \gamma_{i} f(\gamma_{i}) d\gamma_{i} \\
&= E(\gamma_{i})
\end{align*}

\item%c
{\it Find the marginal variance of \(r_{i}\). (Hint: it will be a function of
the prior mean and variance of \(\gamma_{i}\).)}

\vspace{-11pt}
\begin{align*}
Var(r_{i}) &= E\left((r_{i} - E(r_{i}))^{2}\right) \\
&= E\left((r_{i} - E(\gamma_{i}))^{2}\right) \\
&= E\left(r_{i}^2\right) - (E(\gamma_{i}))^{2} \\
&= E\left(E\left(r_{i}^2|\gamma_{i}\right)\right) - (E(\gamma_{i}))^{2} \\
&= \int_{0}^{\infty} E\left(r_{i}^{2}|\gamma_{i}\right)
f(\gamma_{i}) d\gamma_{i} - (E(\gamma_{i}))^{2} \\
&= \int_{0}^{\infty} \left(Var\left(r_{i}|\gamma_{i}\right)
+ \left(E\left(r_{i}|\gamma_{i}\right)\right)^{2}\right)
f(\gamma_{i}) d\gamma_{i} - (E(\gamma_{i}))^{2} \\
&= \int_{0}^{\infty} \left(\frac{\gamma_{i}}{E_{i}}
+ \gamma_{i}^{2}\right)
f(\gamma_{i}) d\gamma_{i} - (E(\gamma_{i}))^{2} \\
&= E\left(\frac{\gamma_{i}}{E_{i}}\right) + E\left(\gamma_{i}^{2}\right)
- (E(\gamma_{i}))^{2} \\
&= \frac{E(\gamma_{i})}{E_{i}} + Var(\gamma_{i})
\end{align*}

\end{enumerate}

\end{enumerate}

\end{document}